---
title: "Fire Skips Analysis"
author: "A. Clason"
date: "July 29, 2020"
output: html_document
---

```{r}
library(data.table)
library(lme4)
```

#Data
```{r}
Plantations <- fread("./Inputs/Plantations.csv")
SitePrep <- fread("./Inputs/SitePrep.csv")
#Leave NA for site prep that hasn't been evaluated
Plant_SP <- merge(Plantations,SitePrep, by=c("OPENING_ID","FireID"),all.x=TRUE) 

```

Fire severity across region
```{r}
#####
```


```{r}
fread()
```


Fire severity in managed vs. non-managed landscapes (probably too much confounding)

- need to have climate and dem etc. here
```{r}

```


# Alternative statistical models
```{r}
# (Silviculture) Spaced + Brushed + SitePrepped + Fertil + Prune + Planted + HarvType 
# (Tree waste treatments)  Pile + Piled Windrow (WINDR, WINDP)+ RPile(road pile) CableKnock|Knock|Mdown(machine knock) |ManualKnock|Snagfalling|Pushover
# (Site and fuels treatments) Soil prep (Disc|Shark|Scalp|Ripping|Brip|Ditching|Mound|Mulch|Patch)
# (Forest Structure) PlantAge + CanopCloVRI + HeightVRI + BA_VRI + DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov
# (Burn of plantation compared to surrounding) DOB + SurCrClos + SurBA + sur_dNBR + Sur-AgeCl-5 + Sur-AgeCl-4 + Sur-AgeCl-3 + Sur-AgeCl-2 + Sur-AgeCl-1
# (Burned previously) TimeSinceWildfire + PileBurn + BroadcastBurn 
# (environment) HeatLoad + TopoPos + Slope 
#(climate) MAT + MAP + CMD

#Extra variables
c("AgeVRI", "HarVRI","NumDenuda", "SurCrClos", "Sur-High", "Sur-Medium", "Sur-AgeCl-2","Sur-Low", "Sur-Unburned",   "Sur-AgeCl-4", "SurM-PineMix", "Sur-AgeCl-1", "SurM-75Pine", "SurM-Pine", "Sur-AgeCl-3", "SurNA-NF", "Sur-AgeCl-5")
#Make sure all factors are factors
Plant_SP[, SitePrepped:=as.factor(SitePrepped)]
Plant_SP[, Fertil:=as.factor(Fertil)]
Plant_SP[, Prune:=as.factor(Prune)]
Plant_SP[, Spaced:=as.factor(Spaced)]
Plant_SP[, Brushed:=as.factor(Brushed)]
Plant_SP[, Planted:=as.factor(Planted)]
Plant_SP[, Reburn:=as.factor(Reburn)]
Plant_SP[, HarvType:=as.factor(HarvType)]
Plant_SP[, FireID:=as.factor(FireID)]
Plant_SP[, HarVRI:=as.factor(HarVRI)]
Plant_SP[, BroadBurn:=as.factor(BroadBurn)]
Plant_SP[, DebrisBurn:=as.factor(DebrisBurn)]
Plant_SP[, DebrisMade:=as.factor(DebrisMade)]
Plant_SP[, DebrisPiled:=as.factor(DebrisPiled)]
Plant_SP[, Soil:=as.factor(Soil)]

#scale and centre continuous variables
Plant_SP[,DecVRICov:=scale(DecVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,ConVRICov:=scale(ConVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,PineVRICov:=scale(PineVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,FirVRICov:=scale(FirVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,SpruVRICov:=scale(SpruVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,DFirVRICov:=scale(DFirVRICov, center=TRUE, scale=TRUE)]
Plant_SP[,BA_VRI:=scale(BA_VRI, center=TRUE, scale=TRUE)]
Plant_SP[,HeightVRI:=scale(HeightVRI, center=TRUE, scale=TRUE)]
Plant_SP[,TotArea:=scale(TotArea, center=TRUE, scale=TRUE)]
Plant_SP[,SurBA:=scale(SurBA, center=TRUE, scale=TRUE)]
Plant_SP[,sur_dNBR:=scale(sur_dNBR, center=TRUE, scale=TRUE)]
Plant_SP[,Md_DOB:=scale(Md_DOB, center=TRUE, scale=TRUE)]
Plant_SP[,DEM_Slop:=scale(DEM_Slop, center=TRUE, scale=TRUE)]
Plant_SP[,DEM_tpi:=scale(DEM_tpi, center=TRUE, scale=TRUE)]
Plant_SP[,DEM_hli:=scale(DEM_hli, center=TRUE, scale=TRUE)]

#Plant_SP[,SurDecPr :=sum(`SurY-Dec`,`SurIM-Dec`,`SurM-Dec`), by="OPENING_ID"]
#Plant_SP[,SurDecPA := ifelse(SurDecPr>0,1,0)] #any deciduous

Plant_SP[,`SurY-Con`:=scale(`SurY-Con`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurIM-Con`:=scale(`SurIM-Con`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurM-Con`:=scale(`SurM-Con`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurY-Dec`:=scale(`SurY-Dec`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurIM-Dec`:=scale(`SurIM-Dec`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurM-Dec`:=scale(`SurM-Dec`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurY-50_50`:=scale(`SurY-50_50`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurIM-50_50`:=scale(`SurIM-50_50`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurM-50_50`:=scale(`SurM-50_50`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurY-75Con`:=scale(`SurY-75Con`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurIM-75Con`:=scale(`SurIM-75Con`, center=TRUE, scale=TRUE)]
Plant_SP[,`SurM-75Con`:=scale(`SurM-75Con`, center=TRUE, scale=TRUE)]
```

## Surrounding forest Type covariates
```{r}

FT_list <- list()
round(cor(Plant_SP[,.(`SurY-Con`,`SurIM-Con`,`SurM-Con`,`SurY-Dec`,`SurIM-Dec`,`SurM-Dec`,`SurY-50_50`,
            `SurIM-50_50`,`SurM-50_50` , `SurY-75Con`,`SurIM-75Con`, `SurM-75Con`)]),2)

#Forest type influence:
FT_list[[1]] <- lmer(dNBR_sc ~ `SurY-Con`+ `SurIM-Con` + `SurM-Con`+
                          `SurY-Dec` + `SurIM-Dec` + `SurM-Dec` + `SurY-50_50`+ `SurIM-50_50`+`SurM-50_50` +
                          `SurY-75Con`+`SurIM-75Con`+ `SurM-75Con` + (1|FireID), 
                        data = Plant_SP)
FT_list[[2]] <- lmer(dNBR_sc ~ `SurY-Con` + (1|FireID), data = Plant_SP)
FT_list[[3]] <- lmer(dNBR_sc ~ `SurIM-Con` + (1|FireID),  data = Plant_SP)
FT_list[[4]] <- lmer(dNBR_sc ~ `SurM-Con` + (1|FireID), data = Plant_SP)#this is the one. Variability, and good reponse
FT_list[[5]] <- lmer(dNBR_sc ~ `SurY-Dec` + (1|FireID),   data = Plant_SP)
FT_list[[6]] <- lmer(dNBR_sc ~ `SurIM-Dec` + (1|FireID), data = Plant_SP)#there's so little of this,
FT_list[[7]] <- lmer(dNBR_sc ~ `SurM-Dec` + (1|FireID),  data = Plant_SP)
FT_list[[8]] <- lmer(dNBR_sc ~ `SurY-50_50` + (1|FireID),data = Plant_SP)
FT_list[[9]] <- lmer(dNBR_sc ~ `SurIM-50_50` + (1|FireID),  data = Plant_SP)
FT_list[[10]] <- lmer(dNBR_sc ~ `SurM-50_50` + (1|FireID), data = Plant_SP)
FT_list[[11]] <- lmer(dNBR_sc ~ `SurY-75Con` + (1|FireID), data = Plant_SP)
FT_list[[12]] <- lmer(dNBR_sc ~ `SurIM-75Con` + (1|FireID), data = Plant_SP)
FT_list[[13]] <- lmer(dNBR_sc ~ `SurM-75Con` + (1|FireID), data = Plant_SP)
lapply(FT_list,FUN = AIC)


FT_list <- list()
FT_list[[1]] <- lmer(dNBR_sc ~ SurBA + (1|FireID), data = Plant_SP)
FT_list[[2]] <- lmer(dNBR_sc ~ SurTotPlArea + (1|FireID), data = Plant_SP)
FT_list[[3]] <- lmer(dNBR_sc ~ sur_dNBR + (1|FireID), data = Plant_SP) #surrounding dNBR obviously best predictor.
FT_list[[4]] <- lmer(dNBR_sc ~ SurMCon + (1|FireID), data = Plant_SP)

FT_list[[5]] <- lmer(dNBR_sc ~ SurBA + SurTotPlArea + SurMCon + (1|FireID), data = Plant_SP)
FT_list[[6]] <- lmer(dNBR_sc ~ sur_dNBR + (1|FireID), data = Plant_SP)

lapply(FT_list,FUN = AIC)

#Change the name of surrounding con
Plant_SP[,SurMCon :=`SurM-Con`]
```


## Testing for correlations between covariates
```{r}
#surrounding forest 
round(cor(Plant_SP[,.(`SurY-Con`,`SurIM-Con`,`SurM-Con`,`SurY-Dec`,`SurIM-Dec`,`SurM-Dec`,`SurY-50_50`,
            `SurIM-50_50`,`SurM-50_50` , `SurY-75Con`,`SurIM-75Con`, `SurM-75Con`)]),2)
#species cover
round(cor(Plant_SP[,.(DecVRICov , ConVRICov , PineVRICov , FirVRICov , SpruVRICov , DFirVRICov)]),2)


```



```{r}

```




## Alternate models
```{r}
Model_list <- list()
#Full model
Model_list[[1]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          PlantAge +DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea + sur_dNBR + 
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No silviculture
Model_list[[2]] <- lmer(dNBR_sc ~ PlantAge +DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea + sur_dNBR + 
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No surrounding forest type
Model_list[[3]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          PlantAge +DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + sur_dNBR + TotArea + DEM_Slop + DEM_tpi + DEM_hli +
                          Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No surrounding dNBR, but surrounding forest type
Model_list[[4]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          PlantAge +DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + SurBA + SurTotPlArea +  
                          SurMCon + TotArea + DEM_Slop + DEM_tpi + DEM_hli +
                          Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No Forest structure
Model_list[[5]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          TotArea + SurBA + SurTotPlArea + sur_dNBR + 
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No Terrain
Model_list[[6]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          PlantAge + DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea + sur_dNBR + 
                          SurMCon + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No PlantAge and no dNBR - full
Model_list[[7]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea +  
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No PlantAge and no dNBR - No silviculture
Model_list[[8]] <- lmer(dNBR_sc ~ DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea +  
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)

#No PlantAge and no dNBR - no forest structure
Model_list[[9]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          TotArea + SurBA + SurTotPlArea +  
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No PlantAge and no dNBR -No surrounding forest type
Model_list[[10]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + DEM_Slop + DEM_tpi + DEM_hli +
                          Md_DOB +  (1|FireID), 
                        data = Plant_SP)
#No PlantAge and no dNBR -No Terrain
Model_list[[11]] <- lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea + 
                          SurMCon + Md_DOB +  (1|FireID), 
                        data = Plant_SP)

lapply(Model_list,FUN = AIC)
summary(Model_list[[7]])

# Factors
ggplot(Plantations,aes(x=Reburn, y=dNBR_sc))+
  geom_boxplot()
# Continuous
ggplot(Plantations,aes(x=sur_dNBR, y=dNBR_sc))+
  geom_point()#+
  geom_smooth(method="gam")

```

##Models for detailed silviculture
There are less than the original number of plantations, as some have been tossed out since the data population project started


Interestingly, when I added the fireID as random effect, brushing and spacing had less influence
```{r}
summary(lm(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType,data = Plant_SP))


Silvi_Model_list <- list()
#Full model
Silvi_Model_list[[1]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled + Soil +
                               Fertil + Prune + Spaced + Brushed + Planted + HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
Silvi_Model_list[[2]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled + Soil +
                               Fertil + Prune + Spaced + Brushed + Planted + HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
Silvi_Model_list[[3]] <- lmer(dNBR_sc ~ DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + 
                                DFirVRICov + BA_VRI + CanopCloVRI +HeightVRI + TotArea + SurBA + 
                                SurTotPlArea + sur_dNBR +DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +
                                `SurY-Con`+ `SurIM-Con` + `SurM-Con`+ `SurY-Dec` + `SurIM-Dec` + 
                                `SurM-Dec` + `SurY-50_50`+ `SurIM-50_50`+`SurM-50_50` + `SurY-75Con` +
                                `SurIM-75Con`+ `SurM-75Con`+ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled +
                                Soil + Fertil + Prune + Spaced + Brushed + Planted + HarvType + PlantAge +
                                 (1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])

lapply(Silvi_Model_list,FUN = AIC)
summary(Silvi_Model_list[[1]])

#No stand management
Silvi_Model_list[[2]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled + Soil + Planted + 
                               HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No harvest
Silvi_Model_list[[3]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled +Fertil + Soil + 
                               Prune + Spaced + Brushed + Planted + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No planting
Silvi_Model_list[[4]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled +Fertil + Soil + 
                               Prune + Spaced + Brushed + HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No soil
Silvi_Model_list[[5]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled + Fertil + Prune +
                               Spaced + Brushed + Planted + HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No piling and burning
Silvi_Model_list[[6]] <- lmer(dNBR_sc ~ BroadBurn + DebrisMade + Soil + Fertil + Prune + Spaced + Brushed + Planted + 
                               HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No debris falling
Silvi_Model_list[[7]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisPiled + Soil + Fertil 
                              + Prune + Spaced + Brushed + Planted + 
                               HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No Broadcast burning
Silvi_Model_list[[8]] <- lmer(dNBR_sc ~ DebrisBurn + DebrisMade + DebrisPiled + Soil + Fertil + Prune + 
                               Spaced + Brushed + Planted + HarvType + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No harvest AND no planting
Silvi_Model_list[[9]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled +Fertil + Soil + 
                               Prune + Spaced + Brushed + PlantAge +(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])
#No stand management, no harvest and no planting
Silvi_Model_list[[10]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + DebrisMade + DebrisPiled + Soil + PlantAge +(1|FireID),
                             data = Plant_SP[!is.na(BroadBurn)])
#No harvest, no planting, no piling and burning management
Silvi_Model_list[[11]] <- lmer(dNBR_sc ~ BroadBurn + DebrisMade + Soil + Fertil + Prune + Spaced + Brushed
                               + PlantAge +(1|FireID),
                              data = Plant_SP[!is.na(BroadBurn)])
#No harvest, no planting, no debris made and no piling and burning
Silvi_Model_list[[12]] <- lmer(dNBR_sc ~ BroadBurn + Soil + Fertil + Prune + Spaced + Brushed
                               + PlantAge +(1|FireID),
                              data = Plant_SP[!is.na(BroadBurn)])

#Just burning treatments
Silvi_Model_list[[13]] <- lmer(dNBR_sc ~ BroadBurn + DebrisBurn + PlantAge +(1|FireID), 
                              data = Plant_SP[!is.na(BroadBurn)])
#Just Debris management
Silvi_Model_list[[14]] <- lmer(dNBR_sc ~ DebrisBurn + DebrisMade + DebrisPiled + PlantAge +(1|FireID), 
                              data = Plant_SP[!is.na(BroadBurn)])
#Just soil treatment
Silvi_Model_list[[15]] <- lmer(dNBR_sc ~ Soil + PlantAge +(1|FireID),
                              data = Plant_SP[!is.na(BroadBurn)])
#Just stand management
Silvi_Model_list[[16]] <- lmer(dNBR_sc ~ Fertil + Prune + Spaced + Brushed + PlantAge +(1|FireID), 
                              data = Plant_SP[!is.na(BroadBurn)])
#Cheating "best model"
Silvi_Model_list[[17]] <- lmer(dNBR_sc ~ DebrisMade  + Soil + Spaced + Brushed + PlantAge+(1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])

Silvi_Model_list[[18]] <- lmer(dNBR_sc ~ DebrisMade  + Soil + Prune + Brushed + PlantAge + (1|FireID), 
                             data = Plant_SP[!is.na(BroadBurn)])


lapply(Silvi_Model_list,FUN = AIC)
summary(Silvi_Model_list[[1]])

###
ggplot(Plant_SP[!is.na(BroadBurn)],aes(x=Spaced, y=dNBR_sc))+
  geom_boxplot()
ggplot(Plant_SP,aes(x=MinTSPF, y=dNBR_sc))+
  geom_point()#+

```


