---
title: "Broadcast Burn Analysis"
author: "A. Clason"
date: "September 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#LOAD LIBRARIES
```{r eval=TRUE,echo=FALSE}
library(data.table)
library(dplyr)
library(raster) #likely don't need this
library(sf) #maybe don't need this
library(stringr)
library(readr)
library(ggplot2)
library(ggpubr)

#set the fires of interest
study_fireTable <- fread("./Inputs/StudyFireList.csv") #all potential fires
FiresOfInterest <- c("C20735","C50744","C20729","C10784","C10970", "R11796",
                       "R11498","G41607", "G51632", "R21721", "C11937",  "R11921")
FiresOfInterest <- c("C20735","C50744","C20729","C10784","C10970", "R11796",
                       "R11498","G41607", "G51632", "R21721", "C11937",  "R11921")

```

##CREATE FIRE DATA TABLES - don't run
```{r}
#source(MakeFireDataTable.R) 
#if you need to actually run the fire table script to generate data tables. Could do it as a function instead where you pass a fireID and create a datatable from that
## 
```

##READ IN PROCESSED DATA
Read in plantations data with site prep treatment methods added
```{r}
dt <- data.table()
Plantations <- data.table()
for(j in 1:length(FiresOfInterest)){
  dt <- fread(paste0("./Outputs/IndividualFires/",FiresOfInterest[j],"_Firedat_SP.csv"))
  dt[, ':='(FireID = FiresOfInterest[j])]
  dt[,.N,by=OPENING_ID]
  dt[,.N]
  Plantations <- rbind(Plantations,dt,fill=TRUE)
}
```
TotArea = total area of the opening in m2
Total_Area = area of fire from others calculations
TotOpenArea calculated here = total area of the opening in ha

#EXTRA DATA PROCESSING STEPS
```{r}
####sCALE dNBR ############ this really should be moved to MakeFireDataTable.R
Plantations[,dNBR_sc:= Mn_dNBR*100] #putting dNBR on the right scale
Plantations[,sur_dNBR_sc:= sur_dNBR*100] #putting dNBR on the right scale
Plantations[,PlantAgeCat := ifelse(PlantAge<11,1,
                                     ifelse(PlantAge<21,2,
                                            ifelse(PlantAge<31,3,
                                                   ifelse(PlantAge<41,4,
                                                          ifelse(PlantAge<51,5,6)))))] #plantation age categories


```

## ANALYSIS
###  Kira's script

Notes -
you can't have treat_area in the same model as broadcast burn because they are highly correlated (not totally right because burn treatment is yes/no), but I think it's more important to compare them as alternative models. 
Same AIC score with treatment area or burn treatment
```{r}
library(lme4)
library(lmerTest)

all<-fread("./Inputs/all_fires2.csv")
str(all)
all[,Burn:=as.factor(Burn)]
all[,Fire_ID:=as.factor(Fire_ID)]
hist(all$mdDNBR_sc)
all[,mdDNBR_sc:=Mn_dNBR*1000]
mean(all[Burn==1,mdDNBR_sc])

#Full model
AIC(lmer(mdDNBR_sc ~ Burn + PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov+ Md_DOB +(1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ Burn +(1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ PlantAge +(1|Fire_ID),data = all))
summary(lm(mdDNBR_sc ~ Burn,data = all))
summary(lm(mdDNBR_sc ~ Burn,data = all[Fire_ID=="Nadina"]))
summary(lm(mdDNBR_sc ~ Burn,data = all[Fire_ID=="Shovel"]))
summary(lm(mdDNBR_sc ~ Burn,data = all[Fire_ID=="Verdun"]))
summary(lm(mdDNBR_sc ~ Burn,data = all[Fire_ID=="Island"]))

summary(lm(mdDNBR_sc ~ PlantAge,data = all))
summary(lm(mdDNBR_sc ~ Burn*PlantAge,data = all))

AIC(lmer(mdDNBR_sc ~ Burn + (1|Fire_ID),data = all)) #random intercepts
AIC(lmer(mdDNBR_sc ~ PlantAge + (1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ Burn+PlantAge + (1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ Burn*PlantAge + (1|Fire_ID),data = all))

AIC(lmer(mdDNBR_sc ~ Burn*PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov + Md_DOB+ (1|Fire_ID),data = all))
summary(lm(mdDNBR_sc ~ Burn+PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov,data = all[Fire_ID=="Nadina"]))
summary(lm(mdDNBR_sc ~ Burn+PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov,data = all[Fire_ID=="Shovel"]))
summary(lm(mdDNBR_sc ~ Burn+PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov,data = all[Fire_ID=="Verdun"]))
summary(lm(mdDNBR_sc ~ Burn+PlantAge + DEM_tpi + BA_VRI + DEM_hli  + cov,data = all[Fire_ID=="Island"]))



AIC(lmer(mdDNBR_sc ~ Burn*PlantAge +  BA_VRI + DEM_hli  + cov + Md_DOB+ (1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ DEM_tpi + BA_VRI + DEM_hli  + cov + Md_DOB+ (1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ DEM_hli  + cov + Md_DOB+ (1|Fire_ID),data = all))
AIC(lmer(mdDNBR_sc ~ Burn*PlantAge + DEM_tpi + BA_VRI + (1|Fire_ID),data = all))

summary(lmer(mdDNBR_sc ~ Burn + (1|Md_DOB),data = all[Fire_ID=="Shovel"]))

AIC(lmer(mdDNBR_sc ~ (1|Fire_ID),data = all))
AIC(glm(mdDNBR_sc ~ 1,data = all))

var.test(all[Burn==0]$mdDNBR_sc,all[Burn==1]$mdDNBR_sc)
t.test(all[Burn==0]$mdDNBR_sc,all[Burn==1]$mdDNBR_sc, var.equal = TRUE)

ggplot()+
  geom_histogram(data=all[Burn==0],aes(x=mdDNBR_sc, fill=Burn))+
  geom_histogram(data=all[Burn==1],aes(x=mdDNBR_sc, fill=Burn))+
  facet_wrap(~Fire_ID)

ggplot(all)+
  geom_boxplot(aes(x=Burn, y=mdDNBR_sc,colour=Fire_ID))

ggplot(all)+
  geom_boxplot(aes(x=Burn, y=PlantAge,colour=Fire_ID))

ggplot(all)+
  geom_boxplot(aes(x=Burn, y=treat_area,colour=Fire_ID))

ggplot(all,aes(x=PlantAge))+
  geom_histogram()+
  facet_wrap(~Burn, nrow=2)

ggplot(all)+
  geom_boxplot(aes(x=Fire_ID, y=mdDNBR_sc))


ggplot(all,aes(x=DEM_tpi, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')
ggplot(all,aes(x=DEM_tpi, y=mdDNBR_sc))+
  geom_point()+
  geom_smooth(method='lm')+
  facet_wrap(~Fire_ID)
ggplot(all,aes(x=Md_DOB, y=mdDNBR_sc))+
  geom_point(aes(colour=Burn,alpha=0.5))+
  #geom_smooth(method='lm')+
  facet_wrap(~Fire_ID)

FullDat <- Plant_SP[FireID=="R21721"|FireID=="R11921"|FireID=="R11498"|FireID=="R11796"]
ggplot(FullDat)+
  geom_point(aes(x=Md_DOB, y=dNBR_sc))+
  geom_smooth(aes(x=Md_DOB, y=dNBR_sc, colour=BroadBurn),method='lm')+
  geom_smooth(aes(x=Md_DOB, y=dNBR_sc, colour=PileBurn),method='lm')+
  geom_smooth(aes(x=Md_DOB, y=dNBR_sc, colour=AnyBurn),method='lm')+
  facet_wrap(~FireID)

mean(FullDat[BroadBurn==0]$dNBR_sc)
mean(FullDat[BroadBurn==1]$dNBR_sc)
t.test(FullDat[BroadBurn==0]$dNBR_sc,FullDat[BroadBurn==1]$dNBR_sc)
t.test(FullDat[PileBurn==0]$dNBR_sc,FullDat[PileBurn==1]$dNBR_sc)
t.test(FullDat[PileBurn==0]$dNBR_sc,FullDat[SpotBurn==1]$dNBR_sc)
t.test(FullDat[PileBurn==0]$dNBR_sc,FullDat[AnyBurn==1]$dNBR_sc)


ggplot(all,aes(x=Md_DOB, y=mdDNBR_sc))+
  geom_point(aes(colour=DEM_tpi,alpha=0.5))+
  #geom_smooth(method='lm')+
  facet_wrap(~Fire_ID)

ggplot(all,aes(x=DEM_hli, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')

ggplot(all,aes(x=BA_VRI, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')

ggplot(all,aes(x=treat_area, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')

ggplot(all,aes(x=Md_DOB, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')

ggplot(all,aes(x=PlantAge, y=mdDNBR_sc,colour=Fire_ID))+
  geom_point()+
  geom_smooth(method='lm')

ggplot(all[Fire_ID=="Island"],aes(x=Md_DOB, y=mdDNBR_sc,colour=Burn))+
  geom_point()
  #geom_smooth(method='lm')

#Full
AIC(lmer(mdDNBR_sc ~ Burn * PlantAge + DEM_tpi + BA_VRI + DEM_hli + cov  +(1|Fire_ID),
             data = all[Fire_ID=="Nadina"|Fire_ID=="Shovel"]))
AIC(lmer(mdDNBR_sc ~ Burn + DEM_tpi + BA_VRI + DEM_hli + cov  +(1|Fire_ID),
             data = all[Fire_ID=="Nadina"|Fire_ID=="Shovel"]))
rsq.glmm(full)
#No random effect on fire
AIC(glm(mdDNBR_sc ~ Burn + PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov , data = all))
#Burn area instead of burn treatment
summary(glm(mdDNBR_sc ~ Burn + PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov , data = all))
#No burn treatment
AIC(glm(mdDNBR_sc ~ PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov, data = all))
#No plantation age 
AIC(glm(mdDNBR_sc ~ DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov, data = all))

#No topography or day of burn (factors important in fire behaviour)
AIC(glm(mdDNBR_sc ~ Burn + PlantAge + BA_VRI + cov, data = all))
#No forest cover
AIC(glm(mdDNBR_sc ~ Burn + PlantAge + DEM_tpi+ DEM_hli + Md_DOB, data = all))
#just age 
AIC(glm(mdDNBR_sc ~ PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov, data = all))

AIC(glm(mdDNBR_sc ~ DEM_tpi, data = all))
AIC(glm(mdDNBR_sc ~ 1 , data = all))
AIC(lmer(mdDNBR_sc ~ 1 + (1|Fire_ID), data = all))
AIC(lmer(mdDNBR_sc ~ Burn + (1|Fire_ID), data = all))



AIC(glm(Mn_dNBR ~ Burn + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov, data = all))



summary(glm(Mn_dNBR ~ Burn + PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov + treat_area, data = all))
summary(glm(Mn_dNBR ~ PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov + treat_area, data = all))


#No DOB, cov, hli
AIC(glm(Mn_dNBR ~ Burn-1 + PlantAge + DEM_tpi + BA_VRI, data = all))
#No DOB, cov, hli
AIC(glm(Mn_dNBR ~ Burn-1 + PlantAge + DEM_tpi + BA_VRI + treat_area, data = all))
#No DOB, cov, hli
AIC(glm(Mn_dNBR ~ PlantAge + DEM_tpi + BA_VRI, data = all))


mod1 <- glm(Mn_dNBR ~ Burn + PlantAge + DEM_tpi + BA_VRI + DEM_hli + Md_DOB + cov + Fire_ID, data = all)
mod2 <- glm(Mn_dNBR ~ Burn + PlantAge + DEM_tpi+BA_VRI, data = all, family = gaussian)
mod3 <- glm(Mn_dNBR ~ Burn + PlantAge + DEM_tpi+BA_VRI+treat_area, data = all, family = gaussian)


mod1 <- glm(Sev~Burn+Age+Dem+Bas+Asp+fire+dob+cov, data = all, family = gaussian)
mod2 <- glm(Sev~Burn+Age+Dem+Bas, data = all, family = gaussian)
mod3 <- glm(Sev~Burn+Age+Dem+Bas+t_area, data = all, family = gaussian)



summary(mod1)
summary(mod2)
summary(mod3)

summary(mod1)$coef
summary(mod2)$coef
summary(mod3)$coef

#assessing deviance
1 - pchisq(deviance(identity),df.residual(identity))

print(identity)

confint(mod1)
coef(mod1)

#looking at the spread of the data
gg_jitter <- ggplot(data = all, aes(x = Sev, y = Dem)) + 
  geom_jitter(width = 0, height = 0.05) +
  ylab("TPI") +
  xlab("Fire severity (dNBR)")

gg_jitter

options(na.action = "na.fail")
dredge(mod1)


get.models(identity)
stepAIC(identity)

#95% CI slope
confint(mod2)
confint(mod3)
#compare the res sum of square
anova(mod2)
anova(mod3)



```

### Alternate models
```{r eval=TRUE,echo=FALSE}
Sev <- "Mn_dNBR"
AgeCov <- "PlantAge"
SiteCov <- "DEM_tpi + DEM_hli" #Data need: extract elevation
BB <- "BroadBurn"
FWI <- "DOB_FWI" #Data need: get FW for DOB for each opening
NumPreds <- 4 #age, site, BB, FWI

#this code is a place holder for when all the covariates are ready to go.
############# Create the alternate model table ##################
Mod_table <- do.call(CJ, replicate(NumPreds, 0:1, FALSE)) #get all possible combinations of 5 variables
setnames(Mod_table,c("TSF","Pltr", "ClimCov","SoilCov","SiteCov")) #removed severity
Mod_table #shows me how many models there should be - must be a better way then this
Mod_table <- Mod_table[,lapply(.SD, as.character)]
for (i in seq_along(Mod_table)) set(Mod_table,i=which(Mod_table[[i]]==1), j=i,
                                    value=as.character(colnames(Mod_table)[i]))
Mod_table[,Mod:=paste(TSF, Pltr, ClimCov, SoilCov, SiteCov, sep = ",")]
mod1 <- Mod_table$Mod
mod1 <- gsub("0,", "", mod1)
mod1 <- gsub(",0", "", mod1)

Model_list<-list()
for(ii in 1:nrow(Mod_table)){
  if(ii==1){
    Model_list[[ii]]<- "1"
  } else {
    Model_list[[ii]] <- paste(unlist(mget(noquote(unlist(strsplit(mod1[ii],","))))),collapse=" + ")
  }
}

##################################################################
######################### Run models #############################
Nmodels <- list()
Gmodels <- list()
for(i in 1:length(Model_list)){
  Nmodels[[i]] <- lapply(CarbonPool1, function(x){
    glm(as.formula(paste(as.name(x),"~",Model_list[[i]])), family=gaussian, data=FR_treatments)
    })
  Gmodels[[i]] <- lapply(CarbonPool2, function(x){
    glm(as.formula(paste(as.name(x),"~",Model_list[[i]])), family="Gamma", data=FR_treatments)
    })
}

#you have to watch the order of the carbon pools
AIC_table <- as.data.table(cbind(Model=c(paste0("Model",seq(1,nrow(Mod_table)))),
                                 MinSoil=0,TotalC=0,FF_C=0,DeadC=0,LiveC=0,mod1))
for(ii in 1:length(Model_list)){
  AIC_table[ii,2] <- round(AIC(Nmodels[[ii]][[1]]),0) 
  AIC_table[ii,3] <- round(AIC(Nmodels[[ii]][[2]]),0)
  AIC_table[ii,4] <- round(AIC(Gmodels[[ii]][[1]]),0) 
  AIC_table[ii,5] <- round(AIC(Gmodels[[ii]][[2]]),0)
  AIC_table[ii,6] <- round(AIC(Gmodels[[ii]][[3]]),0) 
}
AIC_table <- AIC_table[,.(Model,TotalC=as.numeric(TotalC),LiveC=as.numeric(LiveC),
                          DeadC=as.numeric(DeadC),MinSoil=as.numeric(MinSoil),FF_C=as.numeric(FF_C),mod1)]
####################################################################
################## Look at model outputs ###########################

setorderv(AIC_table,c("MinSoil"),1)
AIC_table
```


