---
title: "FourFires-Analysis_spatialRF"
author: "Ingrid Farnell"
date: "14/02/2022"
output: html_document
---

This script runs spatial random forest. 
See https://blasbenito.github.io/spatialRF/ for detailed information

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#------------------------------ Load libraries---------------------------------#
ls <- c("tidyverse", "data.table") # Data Management and Manipulation
ls <- append(ls, c("raster")) # geo comp.
ls <- append(ls, c("pdp", "spatialRF", "randomForestExplainer")) # analysis

# Install if needed -- then load. 
new.packages <- ls[!(ls %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(ls, library, character.only = TRUE)  # load the required packages
rm(ls, new.packages)


#------------------------------ 1. Load data ----------------------------------#
SpatialFilesPath <- "E:/Ingrid/Borealis/BVRCfire"
#SpatialFilesPath <- getwd()
# Set the fires of interest - all 2018 fires with openings
FiresOfInterest <- c("G41607", "G51632", "R11498", "R11796","R11921","R21721")


# Read in the rasters
variable_list <- list.files(paste0(SpatialFilesPath, "/Inputs/Rasters/"),
                            pattern =  paste(FiresOfInterest, sep = "", collapse = "|"), # only import the fires of interest
                            recursive = TRUE,
                            full.names = TRUE)
variable_list <- grep("tif", variable_list, value=TRUE)
# Drop OpenID, None and SitePrepped
variable_list <- grep("OpenID|None|SitePrepped", variable_list, value = TRUE, invert = TRUE)
# For now remove this one because the raster contains only NA's -- Alana to fix and then remove this line
variable_list <- grep("R11498_SpotBurn", variable_list, value = TRUE, invert = TRUE)

variables <- sapply(variable_list, raster)


# Rename the variables 
variable.name <- lapply(str_split(variable_list,"/"), function(x) grep(".tif", x, value=TRUE))
variable.name <- str_split(variable.name, ".tif", simplify = TRUE)[,1]

names(variables) <- variable.name

#ID the names of the categorical rasters
ctg_variables <- c("BEC", "BroadBurn", "Brushed", "DebrisMade", "DebrisPiled", "Fertil", "MechUnk", 
                   "OPENING_ID", "PileBurn", "PlantAge", "Prune", "Soil", "Spaced", 
                   "SpotBurn", "WBurn")
CatRasts <- grep(paste(ctg_variables,sep = "", collapse = "|"),variable.name,value=TRUE)

```

Prepare the data. Resample response and predictor variables to have the same extent and resolution (30 x 30 m) for each of the fires. 
Categorical variables are resampled using nearest neighbourhood and continuous variables are resampled using bilinear resampling. 
Predictor variables must meet the spatial random forest requirements:
  1. Must be free of NA
  2. Columns cannot have 0 variance
  3. Columns must not yield NaN or Inf when scaled

```{r, include = FALSE, warning = FALSE}
# ------------------------------Prepare data-----------------------------------#
#--------------------------2a. Resample rasters and stack----------------------#

# Using base rasters resample response and predictor variables to same extent and resolution
for(i in 1:length(FiresOfInterest)){
  allFireRasts <- variables[grep(FiresOfInterest[i],variables)]
  baseFireRast <- allFireRasts[grep("Base",allFireRasts)][[1]] #index just makes it not a list
  allFireRasts <- allFireRasts[grep("Base",allFireRasts,invert=TRUE)]
  
  # Resample categorical and continuous variables differently
  a <- list()
  for(j in 1:length(allFireRasts)){
    if(names(allFireRasts[[j]]) %in% CatRasts){
      a[[j]] <- resample(allFireRasts[[j]], baseFireRast, method = "ngb")
    } else {
      a[[j]] <- resample(allFireRasts[[j]], baseFireRast, method = "bilinear")
    }
  }
  fireID <- str_extract(names(allFireRasts[1]),FiresOfInterest[i])
  SimpleRastnames <- str_remove(str_remove(names(allFireRasts),FiresOfInterest[i]),"_")
  names(a) <- SimpleRastnames
  #stack the simplified names and assign to fire id rast name
  assign(paste0(fireID,"rasts"), stack(a))
}


#-----------------------------2b. Get sample points----------------------------#
# Create index of raster stacks
RastStacks <- list(G41607rasts, G51632rasts, R11498rasts, R11796rasts, R11921rasts, R21721rasts)

for(i in 1:length(FiresOfInterest)){
  allFireRasts <- variables[grep(FiresOfInterest[i],variables)]
  dNBRFireRast <- allFireRasts[grep("dNBR",allFireRasts)][[1]] #use a raster (doesn't matter which one)
  
  # 270 m grid distance
  b <- aggregate(dNBRFireRast, fact = 9, fun = mean)
  points270 <- rasterToPoints(b, spatial = TRUE) # get sample grid: 1 point/270 m, spatial = TRUE so coordinates are attached
  colnames(points270@data) <- "drop" # make sure to drop this later on(it's a place holder column for points)
  
  # Extract response and predictor values at sample points
  SampledRaster <- raster::extract(RastStacks[[i]], points270, sp = TRUE)
  # Convert to data frame
  dat270 <- as.data.frame(SampledRaster) # hopefully xy = TRUE will attach coordinates, if not do sp = TRUE in above extract line
  
  # Drop rows that don't have an opening ID because we only want to include plantation openings
  dat270 <- dat270 %>% filter(!is.na(OPENING_ID))
  # Drop opening ID column
  dat270 <- subset(dat270, select =-c(OPENING_ID, drop))
  
  # Meet spatial RF requirements
  # 1. Must be free of NA
  dat270 <- dat270[complete.cases(dat270), ] # remove NAs
  
  # 2. Columns cannot have 0 variance
  RemoveZeroVar <- function(dat270) {
    dat270[, !sapply(dat270, function(x) min(x) == max(x))]
  }
  dat270 <- RemoveZeroVar(dat270)
  
  # 3. Columns must not yield NaN or Inf when scaled
  #sum(apply(scale(R11796dat270), 2, is.nan)) 
  #sum(apply(scale(R11796dat270), 2, is.infinite))
  # Find which columns are giving issue
  #sapply(as.data.frame(scale(R21721_270)), function(x)any(is.nan(x)))
  
  # Move response (dNBR) to first column
  dat270 <- dat270 %>% dplyr::select("dNBR", everything())
  fireID <- str_extract(names(allFireRasts[1]),FiresOfInterest[i])
  assign(paste0(fireID,"dat270"), dat270)
  
}

```


```{r, include = FALSE}
# rm(allFireRasts, baseFireRast, dNBRFireRast, G41607rasts, G51632rasts, R11498rasts, R11796rasts, R11921rasts, R21721rasts, points270, SampledRaster, variables, a, b)


```


# G41607 Chutanli Lake
```{r G41607, echo=FALSE}
#----------------------------3a. Set up data-----------------------------------#
# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(G41607dat270)[2:29]

# Coordinates of the cases
xy <- G41607dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(G41607dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 1


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = G41607dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# # Select preds preference order - the ones we want to keep
# preference.order <- c("fwi", "isi", "bui", "dc")
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = G41607dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )

# ** This doesn't seem to be working**


#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = G41607dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
)
spatial.predictors <- spatialRF::get_spatial_predictors(model.spatial)


# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)


# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)


# #----------------------------------6. Tune the model---------------------------#
# Don't do yet, see if spatial model works first
# model.spatial <- rf_tuning(
#   model = model.spatial,
#   xy = xy,
#   repetitions = 30,
#   num.trees = c(500, 1000),
#   mtry = seq(
#     2,
#     length(model.spatial$ranger.arguments$predictor.variable.names), #number of predictors
#     by = 9),
#   min.node.size = c(5, 15),
#   seed = random.seed,
#   verbose = FALSE
# )
# 
# #-----------------------------7. Repeat a model execution----------------------#
# model.spatial.repeat <- spatialRF::rf_repeat(
#   model = model.spatial,
#   repetitions = 30,
#   seed = random.seed,
#   verbose = FALSE
# )
# 
# # Plot variable importance
# spatialRF::plot_importance(
#   model.spatial.repeat,
#   verbose = FALSE
# )
# 
# # Plot partial dependence curves 
# spatialRF::plot_response_curves(
#   model.spatial.repeat,
#   quantiles = 0.5,
#   ncol = 3
# )
# 
# # Model performance
# spatialRF::print_performance(model.spatial.repeat)

```


# G51632 Tezzeron
```{r G51632, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(G51632dat270)[2:23]

# Coordinates of the cases
xy <- G51632dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(G51632dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 2


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = G51632dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = G51632dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )



#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = G51632dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)
```


# R11498 Shovel Lake
Spot Burn was removed because the file is NA - need to fix, but for now I removed spot burn from raster stack
```{r R11498, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(R11498dat270)[2:31]

# Coordinates of the cases
xy <- R11498dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(R11498dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 3


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = R11498dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = R11498dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )



#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = R11498dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)

```


# R11796 Verdun Mountain
```{r R11796, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(R11796dat270)[2:32]

# Coordinates of the cases
xy <- R11796dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(R11796dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 4


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = R11796dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# # Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = R11796dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )
# 


#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = R11796dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)

```


# R11921 Island Lake
```{r R11921, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(R11921dat270)[2:30]

# Coordinates of the cases
xy <- R11921dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(R11921dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 5


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = R11921dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# # Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = R11921dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )
# 
# 

#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = R11921dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)
```


# R21721 Nadina Lake
```{r R21721, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(R21721dat270)[2:33]

# Coordinates of the cases
xy <- R21721dat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(R21721dat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 6


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = R21721dat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# # Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = R21721dat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )
# 


#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = R21721dat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)
```


Entire fire study area (all fires)
# R21721 Nadina Lake
```{r ALL, echo = FALSE}
#----------------------------3a. Set up data-----------------------------------#
# Combine datasets
ALLdat270<- list(G41607dat270, G51632dat270, R11498dat270, R11796dat270, R11921dat270, R21721dat270) %>%
  reduce(full_join)

# If NA make 0 (check first I want to do this)
ALLdat270[is.na(ALLdat270)] <- 0

# Names of response and predictor variables
dependent.variable.name <- "dNBR"
predictor.variable.names <- colnames(ALLdat270)[c(2:29, 32:35)]

# Coordinates of the cases
xy <- ALLdat270[, c("x", "y")]

# Distance matrix
dist <- pointDistance(ALLdat270[, c("x", "y")], lonlat = FALSE)
distance.matrix <- dist

# Distance thresholds (same units as distance matrix)
distance.thresholds <- c(0, 270, 381, 540, 810, 1080, 1350)

# Random seed for reproducibility
random.seed <- 1


#----------------------3b. Assess spatial autocorrelation----------------------#
moran_preds <- spatialRF::plot_training_df_moran(
  data = ALLdat270, 
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100, 
    option = "F", 
    direction = -1),
  point.color = "gray40"
)
moran_preds

#-------------------3c. Reduce autocorrelation in predictors-------------------#
# # Select preds preference order - the ones we want to keep
# preference.order <- c()
# 
# # Assess autocorrelation & drop highly correlated
# predictor.variable.names <- spatialRF::auto_cor(
#   x = ALLdat270[, predictor.variable.names],
#   cor.threshold = 0.75,
#   preference.order = preference.order
# ) %>%
#   spatialRF::auto_vif(
#     vif.threshold = 5,
#     preference.order = preference.order
#   )
# 


#------------------------4. Fit a non-spatial RF model-------------------------#
model.non.spatial <- spatialRF::rf(
  data = ALLdat270,
  dependent.variable.name = dependent.variable.name,
  predictor.variable.names = predictor.variable.names,
  distance.matrix = distance.matrix,
  distance.thresholds = distance.thresholds,
  xy = xy, #not needed by rf, but other functions read it from the model
  seed = random.seed,
  verbose = FALSE
)

# Residuals
spatialRF::plot_residuals_diagnostics(
  model.non.spatial,
  verbose = FALSE
)


#----------------------------4a. Variable importance---------------------------#
# Global variable importance
spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE
)

# Contribution of predictors to model transferability
model.non.spatial <- spatialRF::rf_importance(
  model = model.non.spatial
)

# Local variable importance
local.importance <- spatialRF::get_importance_local(model.non.spatial)

# Partial dependence plots
spatialRF::plot_response_curves(
  model.non.spatial,
  quantiles = 0.5, # other predictors set to 0.5 of quantiles
  ncol = 3
)

#----------------------------4b. Model performance-----------------------------#
spatialRF::print_performance(model.non.spatial)


#-------------------------4c. Spatial cross-validation-------------------------#
model.non.spatial <- spatialRF::rf_evaluate(
  model = model.non.spatial,
  xy = xy,                  #data coordinates
  repetitions = 30,         #number of spatial folds
  training.fraction = 0.75, #training data fraction on each fold
  metrics = "r.squared",
  seed = random.seed,
  verbose = FALSE
)

spatialRF::plot_evaluation(model.non.spatial)
spatialRF::print_evaluation(model.non.spatial)


#---------------------------5. Fit a spatial model-----------------------------#
# Spatial autocorrelation of residuals on non-spatial model
spatialRF::plot_moran(
  model.non.spatial,
  verbose = FALSE
)

# Transform non-spatial model to a spatial model
model.spatial <- spatialRF::rf_spatial(
  model = model.non.spatial,
  method = "mem.moran.sequential", #default method
  verbose = FALSE,
  seed = random.seed
)

# Spatial autocorrelation of residuals on spatial model (what we are trying to reduce from the beginning)
spatialRF::plot_moran(
  model.spatial,
  verbose = FALSE
)

# Variable importance
p1 <- spatialRF::plot_importance(
  model.non.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Non-spatial model")

p2 <- spatialRF::plot_importance(
  model.spatial,
  verbose = FALSE) +
  ggplot2::ggtitle("Spatial model")

p1 | p2

# Top 10 spatial predictors
head(model.spatial$importance$per.variable, n = 10)
```

