---
title: "PlantationSkils_DataPrep"
author: "A. Clason"
date: "March 31, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load libraries
```{r eval=TRUE,echo=FALSE}
library(data.table)
library(dplyr)
library(raster)
library(sf)
library(stringr)
library(readr)
library(ggplot2)
library(ggpubr)

#Study fires
study_fireTable <- fread("./Inputs/StudyFireList.csv")
Fire_shortList <- c("K20637","C20735","C50647","C50744","C20729","C10784","C10970", "R11796",
                    "R11498","G41607", "G51632", "R21721", "C11937",  "R11921")
ResultsPolyOverlap <- fread("./Outputs/IndividualFires/ResultsPolyOverlap.csv")
dNBR_imageryDates <- fread("./Inputs/dNBR_dates.csv")
SitePrepGroups <- fread("./Inputs/SitePrep_TypeMethods.csv")

```

##Spatial Files - only run if running MakeFireDataTable.R

Most of these files need to be loaded here. There are a few (DOB and dNBR) that are loaded fire by fire in the MakeFireDataTable.R script 

If you try and load in the spatial files that were saved from MakeFireDataTable.R, they might not be perfect, due to issues with write_sf function
```{r}
#source(MakeFireDataTable.R) 
#if you need to actually run the fire table script to generate data tables. Could do it as a function instead where you pass a fireID and create a datatable from that

##### Fire perimeters #####
fire_perimeters <- read_sf("F:/Spatial Data/FCI_Fire_Regen_spatialfiles/Historical wildfire perimeters/BC Wildfire Historical Fire Perimeters.shp",quiet=TRUE)
fire_per_sel <- fire_perimeters %>%
  dplyr::select(FIRE_NUMBE,FIRE_YEAR,FIRE_CAUSE)

Study_fire_perimeters <- as.data.table(fire_per_sel)[FIRE_NUMBE %in% c(study_fireTable[,FireID])]

##### Severity mapping (all 2015-2018) #####
severity <- read_sf("F:/Spatial Data/FCI_Fire_Regen_spatialfiles/Burn severity 2015_2018/BurnSeverityRating_2018.shp", quiet=T)
severity_sel <- severity %>%
  dplyr::select(FIRE_NUMBE,FIRE_YEAR,BURN_SEVER)

##### dNBR #####
dNBR_list <- list()
dNBR_list[[1]] <- raster("E:/Fires_dNBR/dNBR_imagery/C10784_Plateau_Complex_dNBR.tif")
dNBR_list[[2]] <- raster("E:/Fires_dNBR/dNBR_imagery/C10970_Kluskoil_Lake_dNBR.tif")
dNBR_list[[3]] <- raster("E:/Fires_dNBR/dNBR_imagery/C11837_Shag_Creek_dNBR.tif")
dNBR_list[[4]] <- raster("E:/Fires_dNBR/dNBR_imagery/C11937_North_Baezaeko_dNBR.tif")
dNBR_list[[5]] <- raster("E:/Fires_dNBR/dNBR_imagery/C20279_Wildwood_dNBR.tif")
dNBR_list[[6]] <- raster("E:/Fires_dNBR/dNBR_imagery/C50647_Hanceville_Riske_Creek_dNBR.tif")
dNBR_list[[7]] <- raster("E:/Fires_dNBR/dNBR_imagery/C50744_Kleena_Kleene_dNBR.tif")
dNBR_list[[8]] <- raster("E:/Fires_dNBR/dNBR_imagery/G04226_Lejac_dNBR.tif")
dNBR_list[[9]] <- raster("E:/Fires_dNBR/dNBR_imagery/G10037_BobtailLk_dNBR.tif")
dNBR_list[[10]] <- raster("E:/Fires_dNBR/dNBR_imagery/G41607_Chutanli_dNBR.tif")
dNBR_list[[11]] <- raster("E:/Fires_dNBR/dNBR_imagery/G51632_Tezzeron_dNBR.tif")
dNBR_list[[12]] <- raster("E:/Fires_dNBR/dNBR_imagery/K20637_Elephant_Hill_dNBR.tif")
dNBR_list[[13]] <- raster("E:/Fires_dNBR/dNBR_imagery/dNBR_imagery/R11498_Shovel_Lake_dNBRS2.tif")
dNBR_list[[14]] <- raster("E:/Fires_dNBR/dNBR_imagery/R11796_Verdun_dNBR.tif")
dNBR_list[[15]] <- raster("E:/Fires_dNBR/dNBR_imagery/R11921_Island_Lake_dNBR.tif")
dNBR_list[[16]] <- raster("E:/Fires_dNBR/dNBR_imagery/R12068_PondosyBay_dNBR.tif")
dNBR_list[[17]] <- raster("E:/Fires_dNBR/dNBR_imagery/R12594_Baldface_Mtn_dNBR.tif")
dNBR_list[[18]] <- raster("E:/Fires_dNBR/dNBR_imagery/R21721_Nadina_dNBR.tif")
dNBR_list[[19]] <- raster("E:/Fires_dNBR/dNBR_imagery/R91947_AlkaliLk_dNBR.tif")
dNBR_list[[20]] <- raster("E:/Fires_dNBR/dNBR_imagery/VA1787_RamseyCrk_dNBR.tif")
dNBR_list[[21]] <- raster("E:/Fires_dNBR/dNBR_imagery/VA1964_DeanRiver_dNBR.tif")

###### Topography ######
DEMslope <- raster("./Inputs/Rasters/DEMslope.tif")
DEMaspect <- raster("./Inputs/Rasters/DEMaspect.tif")
DEMtpi <- raster("./Inputs/Rasters/DEMtpi.tif")
DEMhli <- raster("./Inputs/Rasters/DEMhli.tif")


##### Results #####
#could include Generalize- these describe the BEC subzone/site series
Results_All <- read_sf("F:/Spatial Data/FCI_Fire_Regen_spatialfiles/RESULTS/RESULTS_FirePerimeter_Intersect.shp",
                       quiet=T)
Results_sel <- Results_All %>%
    dplyr::select(OPENING_ID,OPENING_ST,APPROVE_DA,DISTURBANC,DISTURBA_1,DENUDATION, DENUDATI_1, 
                  DENUDATI_2, DENUDATI_3,
                  DENUDATI_4,DENUDATI_5,DENUDATI_6,DENUDATI_7, DENUDATI_8, DENUDATI_9, DENUDAT_10, SITE_PREP_,
                  SITE_PREP1, SITE_PRE_1, SITE_PRE_2, SITE_PRE_3, SITE_PRE_4 ,SITE_PRE_5, PLANTING_1,PLANTING_2,
                  PLANTING_3, PLANTING_4, PLANTING_5, PLANTING_6, PLANTING_C ,BRUSHING_T,BRUSHING_1, BRUSHING_C,
                  BRUSHING_2 ,SPACING_TR, SPACING_CO ,SPACING__1, FERTILIZAT,FERTILIZ_1, FERTILIZ_2, PRUNING_TR,
                  PRUNING_CO ,PRUNING__1,SLOPE,ASPECT)

VRI_study <- read_sf("F:/Spatial Data/Fire/ForestManage_Fire/VRI2016_StudyFires.shp")

##### BEC database ##### 
#BEC <- read_sf("F:/Spatial Data/BEC/Polygons/BECpolygons.shp")


```

## 

TotArea = total area of the opening in m2
Total_Area = area of fire from others calculations
TotOpenArea calculated here = total area of the opening in ha
```{r}
dt <- data.table()
Plantations <- data.table()
for(j in 2:length(Fire_shortList)){
  dt <- fread(paste0("./Outputs/IndividualFires/",Fire_shortList[j],"_Firedat.csv"))
  dt[, ':='(FireID = Fire_shortList[j])]
  dt[,.N,by=OPENING_ID]
  dt[,.N]
  Plantations <- rbind(Plantations,dt,fill=TRUE)
}
Plantations <- merge(Plantations,study_fireTable, by="FireID")
############ sCALE dNBR ############
Plantations[,dNBR_sc:= Mn_dNBR*100]
Plantations[,sur_dNBR_sc:= sur_dNBR*100]
Plantations[,PlantAgeCat := ifelse(PlantAge<11,1,
                                   ifelse(PlantAge<21,2,
                                          ifelse(PlantAge<31,3,
                                                 ifelse(PlantAge<41,4,
                                                        ifelse(PlantAge<51,5,6)))))]

#change every NA to a 0:
#for (i in seq_along(Plantations)) set(Plantations, i=which(is.na(Plantations[[i]])), j=i, value=0)

########## remove plantations that are too young ############
preDNBR <- dNBR_imageryDates[PrePost_Fire=="Pre-fire"]
postDNBR <- dNBR_imageryDates[PrePost_Fire=="Post-fire"]
preDNBR[, PreDate:=as.Date(ImageDate2,format="%d/%m/%Y")]
postDNBR[, PostDate:=as.Date(ImageDate2,format="%d/%m/%Y")]
FireMinDate <- preDNBR[,.(OldestDNBR = min(PreDate)),by="FireNumber"] #what is the oldest date for pre-fire imagery for a fire
range(format(FireMinDate[,OldestDNBR],"%m")) #when was preimagery taken
range(format(postDNBR[,PostDate],"%m")) #when was postimagery taken
FireStartdNBR <- merge(study_fireTable[FireID %in% Fire_shortList,.(FireID,FireName,StartDate)],
                       FireMinDate, by.x="FireID", by.y="FireNumber")
FireStartdNBR[,MinPlantAge := as.numeric(format(as.Date(StartDate,format="%d/%m/%Y"),"%Y"))-
                as.numeric(format(OldestDNBR,"%Y"))]
Plantations <- Plantations[PlantAge>=max(FireStartdNBR[,MinPlantAge])] #using 3 myear minimum for now
##############################################################

### Surrounding forest 
PrSurAgeFT <- melt(Plantations, id.vars= "OPENING_ID",
                   measure.vars =  c("SurY-Dec","SurY-50_50", "SurY-75Con", "SurY-Con", "SurIM-Con","SurIM-50_50",
                                     "SurIM-75Con","SurIM-Dec","SurM-50_50", "SurM-75Con","SurM-Con",
                                     "SurM-Dec","SurNA-NF"), 
                   variable.name = "SurAgeFT",
                   value.name = "SurPr_AgeFT")
PrMatPine <- melt(Plantations, id.vars= "OPENING_ID",
                   measure.vars =  c("SurM-PineMix","SurM-75Pine", "SurM-Pine"), 
                   variable.name = "SurMatPine",
                   value.name = "SurPr_AgePine")
PrSurBS <- melt(Plantations, id.vars="OPENING_ID",
                measure.vars = c("Sur-High","Sur-Medium","Sur-Low","Sur-Unburned"),
                variable.name = "SurBS",
                value.name = "SurPr_BS")
PrSurPlaCl <- melt(Plantations, id.vars="OPENING_ID",
                measure.vars = c("Sur-AgeCl-1","Sur-AgeCl-2","Sur-AgeCl-3","Sur-AgeCl-4","Sur-AgeCl-5"),
                variable.name = "SurPlaCl",
                value.name = "SurPr_PlaCl")

Plantations_SurBS <- merge(Plantations,PrSurBS, by="OPENING_ID")
Plantations_SurAgeFT <- merge(Plantations,PrSurAgeFT, by="OPENING_ID")
Plantations_SurMatPine <- merge(Plantations,PrMatPine, by="OPENING_ID")
Plantations_SurPlaCl <- merge(Plantations,PrSurPlaCl, by="OPENING_ID")

#add up all the area in plantations surrounding a plantation (no age)
Plantations <- merge(Plantations,PrSurPlaCl[,.(SurTotPlArea =sum(na.omit(SurPr_PlaCl))),by="OPENING_ID"], by="OPENING_ID")
Plantations[,SurTotPlArea:=ifelse(SurTotPlArea<1,SurTotPlArea,1)]

#Change to years since previous fire
cols <- colnames(Plantations)[grep("^*[0-9]",colnames(Plantations))]
PrevFire <- melt(Plantations, id.vars= "OPENING_ID",
                   measure.vars =  cols, 
                   variable.name = "PrevFireYr",
                   value.name = "PrevFirePr")
PrevFire <- merge(Plantations[,.(OPENING_ID,StartDate)],PrevFire, by="OPENING_ID")
PrevFire[,StartYear := as.numeric(format(as.Date(StartDate,format="%d/%m/%Y"),"%Y"))]
PrevFire[,PrevFireYr := as.numeric(format(as.Date(PrevFireYr,format="%Y"),"%Y"))]
PrevFire[,TSPreF := ifelse(is.na(PrevFirePr),NA, StartYear - PrevFireYr)]
#the calculation above says any proportion of previous fire counts

#get rid of the year columns:
Plantations[,(cols):=NULL,with=FALSE]
Plantations <- merge(Plantations,
                     PrevFire[!is.na(PrevFirePr),.(NumPrevFires=.N, MinTSPF=min(TSPreF)),by=OPENING_ID],
                     by="OPENING_ID",all.x=TRUE)
#change all NAs to 0
for (i in seq_along(Plantations)) set(Plantations, i=which(is.na(Plantations[[i]])), j=i, value=0)
#write.csv(Plantations, "./Inputs/Plantations.csv",row.names = FALSE)

############# Combining polygons that have been re-numbered #############
#PlantReNum <- ResultsPolyOverlap[PropResOverlap==1]
#overlapFull <- Results_sel %>%
 # filter(OPENING_ID %in% c(PlantReNum$OPENING_ID,PlantReNum$OPENING_ID.1))
#write_sf(overlapFull,"overlapFull.shp")
#Plantations[OPENING_ID %in% PlantReNum$OPENING_ID.1]
#Plantations[OPENING_ID %in% c(-512650000,1658741)]
#Plantations[OPENING_ID %in% PlantReNum[1]]
#PlantOver <- merge(Plantations,ResultsPolyOverlap, by="OPENING_ID", all.y=T)
#overlapPlantations[PropResOverlap>0.9] #polygons with 100% overlap (30)

####################################
#look at relationships
ggplot(Plantations,aes(x=PlantAge, y=dNBR_sc))+
  geom_point()+
  geom_smooth(method="gam")

ggplot(Plantations,aes(x=PlantAge, y=dNBR_sc))+
  geom_point()+
  geom_smooth(method="lm")

ggplot(Plantations,aes(x=SurTotPlArea, y=dNBR_sc))+
  geom_point()+
  geom_smooth(method="lm")

ggplot(Plantations,aes(x=dNBR_sc, y=sur_dNBR_sc))+
  geom_point(aes(colour=factor(PlantAgeCat)))+
  geom_smooth(method="lm", formula = y ~ x-1)




ggplot(Plantations_melt[SurBS=="Sur-High"],aes(x=High, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm", formula=y ~ x-1)
ggplot(Plantations[SurBS=="Sur-Medium"],aes(x=Medium, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")
ggplot(Plantations[SurBS=="Sur-Low"],aes(x=Low, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")
ggplot(Plantations[SurBS=="Sur-Unburned"],aes(x=Unburned, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")

#find the skips?
ggplot(Plantations_melt[SurBS=="Sur-High"],aes(x=Unburned, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")
ggplot(Plantations_melt[SurBS=="Sur-Medium"],aes(x=Unburned, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")
ggplot(Plantations_melt[SurBS=="Sur-Low"],aes(x=Unburned, y=SurPr_BS))+
  geom_point()+
  geom_smooth(method="lm")

#looking at relationship to surounding forest
ggplot(Plantations_SurAgeFT[SurAgeFT=="SurY-Con"|SurAgeFT=="SurIM-Con"|SurAgeFT=="SurM-Con"],
       aes(x=SurPr_AgeFT,y=dNBR_sc,colour=SurAgeFT))+
  geom_point()+
  geom_smooth(method="lm", formula=y~x)
ggplot(Plantations_SurBS,
       aes(x=SurPr_BS,y=dNBR_sc,colour=SurBS))+
  geom_point()+
  geom_smooth(method="lm", formula=y~x-1)

ggplot(Plantations_SurPlaCl,
       aes(x=SurPr_PlaCl,y=dNBR_sc,colour=SurPlaCl))+
  geom_point()+
  geom_smooth(method="lm", formula=y~x-1)


```


```{r }
##### ADD DETAILED SP AND PLANTS #####
dt <- data.table()
SitePrep <- data.table()
for(i in 2:14){
  dt <- fread(paste0("./Outputs/IndividualFires/",Fire_shortList[i],"_Firedat_SitePrep_MethAdds.csv"),
                     na.strings=c("","NA","<NA>"))
  
  cols <-c("SITE_PREP_","SITE_PRE_2",colnames(dt)[grepl("Type",colnames(dt))])
  cols <- cols[grepl("SITE",cols)]
  dt[, (cols):=lapply(.SD, as.factor),.SDcols=cols]
  SitePrepType_melt <- melt(dt, id.vars = c("OPENING_ID"),
                      measure.vars = cols,
                      variable.name = "SP",
                      value.name = "SP_type",value.factor=TRUE)
 
  #Site prep method (pburn, trail, knock down etc.)
  cols <- colnames(dt)[grepl("Meth",colnames(dt))]
  cols <- cols[grepl("SITE",cols)]
  dt[, (cols):=lapply(.SD, as.factor),.SDcols=cols]
  SitePrepMeth_melt <- melt(dt, id.vars = c("OPENING_ID"),
                        measure.vars = cols,
                        variable.name = "SP_",
                        value.name = "SP_Method",value.factor=TRUE)
  #Site prep date
  cols <-c("SITE_PRE_1","SITE_PRE_4",colnames(dt)[grepl("Date",colnames(dt))])
  cols <- cols[grepl("SITE",cols)]
  dt[, (cols):=lapply(.SD,function(x) as.numeric(format(as.Date(x,tryFormats=c("%d/%m/%Y",
                                                                               "%d-%m-%Y",
                                                                               "%Y/%m/%d",
                                                                               "%Y-%m-%d")),"%Y"))),.SDcols=cols]
  SitePrepDate_melt <- melt(dt, id.vars = c("OPENING_ID"),
                        measure.vars = cols,
                        variable.name = "SP_3",
                        value.name = "SP_Date")
  
  
  SitePr <- SitePrepType_melt[,.(OPENING_ID,SP_type,
                                   SP_Method = SitePrepMeth_melt[,SP_Method],
                                   SP_Date = SitePrepDate_melt[,SP_Date])]
  SitePr <- SitePr[, SP_type_meth:= paste0(SP_type,"_",SP_Method)]
  SitePr[, ':='(FireID = Fire_shortList[i])]
  SitePrep <- rbind(SitePrep,SitePr,fill=TRUE)
}
unique(SitePrep$SP_type_meth)
table(SitePrep$SP_type_meth) #WATCH THAT THERE ARE NO NEW COMBINATIONS WITH NEW DATA ENTERED!!
SitePrep[SP_type_meth=="BU_LRIP"]

SitePrep[,SPgrUse:=1]
#SitePrep <- SitePrep[!is.na(SP_type)&!is.na(SP_Method)]
SitePrep[,.N,by="SP_type_meth"]
SitePrep[SP_type_meth=="ME_WINDP"]
#Bring in the grouping variable - only site preps of interest are kept
SitePrep <- merge(SitePrep,SitePrepGroups,by.x="SP_type_meth", by.y="Type_Method", all.x=TRUE)
SitePrep[!is.na(GroupName)]
SitePrepCast <- dcast(SitePrep[!is.na(GroupName)], OPENING_ID~GroupName, value.var ="SPgrUse",fun.aggregate=sum)
for (i in seq_along(SitePrepCast)) set(SitePrepCast, i=which(SitePrepCast[[i]]>0), j=i, value=1)

#merge with the main dataset and make sure any opening without a value has a zero.
SitePrep2 <- merge(SitePrepCast,SitePrep[,.(OPENING_ID,FireID)], by="OPENING_ID", all.y=TRUE)
for (i in seq_along(SitePrep2)) set(SitePrep2, i=which(is.na(SitePrep2[[i]])), j=i, value=0)
SitePrep2<- unique(SitePrep2)
SitePrep2[which(duplicated(SitePrep2[,OPENING_ID]))]
#write.csv(SitePrep,"./Inputs/SitePrep.csv",row.names = FALSE)

Plant_SP <- merge(Plantations,SitePrep2, by=c("OPENING_ID","FireID"), all.x=TRUE)
#missing results data from 50 openings in C50647 - would take 30 minutes to do
Plant_SP <- Plant_SP[!is.na(BroadBurn),] #need to update this when add those openings
cols <- c("BroadBurn", "DebrisBurn","DebrisMade","DebrisPiled","DebrisPiledBurn","Soil","WildBurn","MechUnk")
Plant_SP[,(cols):=lapply(.SD, as.factor),.SDcols=cols] #make sure site prep methods are factors
#need to define debris management:

Plant_SP[DebrisPiled=="1" & DebrisBurn =="1"|DebrisPiledBurn=="1",
         DPile_DBU:= as.factor("1")] #if Debris piled and burned
Plant_SP[DebrisPiled=="1" & DebrisBurn =="0",
         DPileNo_BU:= as.factor("1")] #if Debris made and no burning - maybe include wildfire

#Made but not burned and no soil == Fuels left
Plant_SP[DebrisMade== "1" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="0"|
           DebrisPiled== "1" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="0", 
         FuelLeft:= as.factor("1")]

Plant_SP[DebrisMade=="1" & DebrisPiled=="1" & DebrisBurn=="1" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="0"| 
          DebrisMade=="1" & DebrisPiled=="0" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="1" & Soil=="0"|
           DebrisMade=="1" & DebrisPiled=="0" & DebrisBurn=="0" & DebrisPiledBurn=="1" & BroadBurn=="0" & Soil=="0"|
           DebrisMade=="1" & DebrisPiled=="0" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="1"|
           DebrisMade=="0" & DebrisPiled=="1" & DebrisBurn=="1" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="0"|
           DebrisMade=="0" & DebrisPiled=="0" & DebrisBurn=="1" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="0"|
           DebrisMade=="0" & DebrisPiled=="1" & DebrisBurn=="1" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="1"|
           DebrisMade=="0" & DebrisPiled=="0" & DebrisBurn=="0" & DebrisPiledBurn=="1" & BroadBurn=="0" & Soil=="1"|
           DebrisMade=="0" & DebrisPiled=="1" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="0" & Soil=="1"|
           DebrisMade=="0" & DebrisPiled=="1" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="1" & Soil=="0"|
           DebrisMade=="0" & DebrisPiled=="1" & DebrisBurn=="0" & DebrisPiledBurn=="0" & BroadBurn=="1" & Soil=="1",
         FuelGone:= as.factor("1")]

#then do glms that compare all this treatment info together, or just soil on its own?
summary(lmer(dNBR_sc ~ SitePrepped + Fertil + Prune + Spaced + Brushed + Planted + HarvType + 
                          PlantAge +DecVRICov + ConVRICov + PineVRICov + FirVRICov + SpruVRICov + DFirVRICov + 
                          BA_VRI + CanopCloVRI + HeightVRI + TotArea + SurBA + SurTotPlArea + sur_dNBR + 
                          SurMCon + DEM_Slop + DEM_tpi + DEM_hli + Md_DOB +  (1|FireID), 
                        data = Plant_SP))


Plant_SP[DebrisPiled== "1" & DebrisBurn=="1" | 
           DebrisPiled== "1" & DebrisPiledBurn=="1" | 
           DebrisPiledBurn=="1" |
           DebrisPiled== "1" & BroadBurn=="1", DPile_DBU:= as.factor("1")] #these will also include debris made
Plant_SP[DMadeNo_BU==1]

#I want to know what the treatments were that would lead to a debris being made but not burned
Plant_SP[BroadBurn==1]

#the subset of sites that had debris piled:
Plant_SP[DPile_DBU==1,.N] #12 piled and not burned, 80 piled and burned - that seems low
Plant_SP[DPileNo_BU=="1"|DPile_DBU=="1"]
ggplot(Plant_SP, aes(x=FuelGone,y=Mn_dNBR))+
  geom_boxplot()
ggplot(Plant_SP, aes(x=Soil,y=Mn_dNBR))+
  geom_boxplot()

a <-ggplot(Plant_SP, aes(x=BroadBurn,y=Mn_dNBR))+
  geom_boxplot()
b <-ggplot(Plant_SP, aes(x=DebrisBurn,y=Mn_dNBR))+
  geom_boxplot()
c <-ggplot(Plant_SP, aes(x=DebrisMade,y=Mn_dNBR))+
  geom_boxplot()
d <-ggplot(Plant_SP, aes(x=DebrisPiled,y=Mn_dNBR))+
  geom_boxplot()
e <-ggplot(Plant_SP, aes(x=Soil,y=Mn_dNBR))+
  geom_boxplot()
ggarrange(a,b,c,d,e, ncol=3, nrow=2)

summary(glm(Mn_dNBR ~ Soil, data=Plant_SP))
summary(glm(Mn_dNBR ~ DebrisPiled, data=Plant_SP))
summary(glm(Mn_dNBR ~ DebrisMade, data=Plant_SP))
summary(glm(Mn_dNBR ~ DebrisBurn, data=Plant_SP))
summary(glm(Mn_dNBR ~ BroadBurn, data=Plant_SP)) #so unbalanced
t.test(Plant_SP[Soil==0]$Mn_dNBR, Plant_SP[Soil==1]$Mn_dNBR, var.equal = FALSE)
t.test(Plant_SP[DebrisPiled==0]$Mn_dNBR, Plant_SP[DebrisPiled==1]$Mn_dNBR, var.equal = FALSE)
t.test(Plant_SP[DebrisMade==0]$Mn_dNBR, Plant_SP[DebrisMade==1]$Mn_dNBR, var.equal = FALSE)
t.test(Plant_SP[DebrisBurn==0]$Mn_dNBR, Plant_SP[DebrisBurn==1]$Mn_dNBR, var.equal = FALSE)
t.test(Plant_SP[BroadBurn==0]$Mn_dNBR, Plant_SP[BroadBurn==1]$Mn_dNBR, var.equal = FALSE)

ggplot(SitePrep[!is.na(GroupName)],aes(x=SP_Date, fill=GroupName))+
  geom_histogram(position = "stack")
ggplot(SitePrep[GroupName=="BroadBurn"],aes(x=SP_Date, fill=GroupName))+
  geom_histogram()+
  xlim(c(1960,2020))+
  theme(text=element_text(size=20)) 

ggplot(SitePrep[GroupName=="DebrisMade"|GroupName=="DebrisBurn"],aes(x=SP_Date, fill=GroupName))+
  geom_histogram()+
  xlim(c(1960,2020))+
  theme(text=element_text(size=20))
  
ggplot(SitePrep[GroupName=="Soil"],aes(x=SP_Date, fill=SP_Method))+
  geom_histogram()+
  xlim(c(1960,2020)) +
  theme(text=element_text(size=20))

#############################################################

#Planting Species - not done. Need to decide what is important
PlantType_melt <- melt(FireDetRes, id.vars = c("OPENING_ID"),
                      measure.vars = c("PLANTING_1","PLANTING_4",
                                       "PLANT_3_Type",
                                       "PLANT_4_Type",
                                       "PLANT_5_Type",
                                       "PLANT_6_Type"),
                      variable.name = "PL",
                      value.name = "PL_type")

#Planting Density
PlantSp_melt <- melt(FireDetRes, id.vars = c("OPENING_ID"),
                      measure.vars = c("PLANT_1_SP1",
                                       "PLANT_1_SP1",
                                       "PLANT_1_SP1",
                                       "PLANT_1_SP1",
                                       "PLANT_1_SP1",
                                       "PLANT_1_SP1"),
                      variable.name = "SP_",
                      value.name = "SP_Method")
#Planting Date
SitePrepMeth_melt <- melt(FireDetRes, id.vars = c("OPENING_ID"),
                      measure.vars = c("SITEPr_1_Meth",
                                       "SITEPr_2_Meth",
                                       "SITEPr_3_Meth",
                                       "SITEPr_4_Meth",
                                       "SITEPr_5_Meth",
                                       "SITEPr_6_Meth"),
                      variable.name = "SP_",
                      value.name = "SP_Method")



```


```{r}
###### SUMMARIES #######
Plantations[,TotOpenArea:=TotArea/10000] #area of the RESULTS opening
Plantations[,.N,by=OPENING_ID]
Plantations[,.N]
#check to make sure all the plantations are in fact plantations:
Plantations[,.N,c("DENUDATION","DENUDATI_5")]
Plantations[DENUDATION =="B" & is.na(DENUDATI_5)] #sites just with Burn call have site prep and planting. Good.
Plantations[DENUDATION =="W" & is.na(DENUDATI_5)] #same with windthrow
Plantations[DENUDATION =="L" & DENUDATI_5=="D"]
###########

```


### 
```{r}
ggplot(PlantationsAge, aes(FireID, PlantAge))+
  geom_boxplot()

## graphs of proportion burned by plantation age
a <- ggplot(PlantationsAge,aes(y=Unburned, x=PlantAge))+
  geom_point()+
  geom_smooth(method = "gam")+
  ylim(0,1)
b <- ggplot(PlantationsAge,aes(y=Low, x=PlantAge))+
  geom_point()+
  geom_smooth(method = "gam")+
  ylim(0,1)
c <- ggplot(PlantationsAge,aes(y=Medium, x=PlantAge))+
  geom_point()+
  geom_smooth(method = "gam")+
  ylim(0,1)
d <- ggplot(PlantationsAge,aes(y=High, x=PlantAge))+
  geom_point()+
  geom_smooth(method = "gam")+
  ylim(0,1)
ggarrange(a,b,c,d)
```

## Treatments
Here I'm adding the site treatments associated with openings 
```{r}
###### TREATMENTS (0,1) ########
PlantationsAge[,Spaced := ifelse(SPACING__1==0,0,1)]
PlantationsAge[,Brushed := ifelse(BRUSHING_1==0,0,1)]
PlantationsAge[,SitePrepped := ifelse(SITE_PRE_5==0,0,1)]
PlantationsAge[,Fertil := ifelse(FERTILIZ_2==0,0,1)]
PlantationsAge[,Prune := ifelse(PRUNING__1==0,0,1)]

####### Making sure the severity is 0 if there was none (not NA)
PlantationsAge[,High:= ifelse(is.na(High),0,High)]
PlantationsAge[,Medium:= ifelse(is.na(Medium),0,Medium)]
PlantationsAge[,Low:= ifelse(is.na(Low),0,Low)]
PlantationsAge[,Unburned:= ifelse(is.na(Unburned),0,Unburned)]
############################

#summary information about openings in these study fires:
PlantationsAge[,.N,by=FireID]
#elephant hill has way more than others
PlantationsAge[,max(PlantAge),by=FireID]
PlantationsAge[is.na(PLANTING_1) & is.na(PLANTING_4),.N,by=FireID] #number of unplanted openings
PlantationsAge[is.na(PLANTING_1) & is.na(PLANTING_4),min(PlantAge),by=FireID] #age of unplanted openings

write.csv(PlantationsAge,"./Outputs/PlantationsAge.csv", row.names=FALSE)


```


## Fire-level summaries
### Fire Severity in plantations vs. Total Fire
- Do we want to look at plantations vs. not plantations, or is it plantations vs. total fire (how it is now)
```{r}
#total area of the fire
fires <- fire_per_sel %>% filter(FIRE_NUMBE %in% Fire_shortList)
fires$TotFireArea <- st_area(fires)
firesDT <- as.data.table(fires)
firesDT[,TotFireArea := unclass(TotFireArea)/10000]
firesDTmerge <- firesDT[,.(FIRE_NUMBE,FIRE_YEAR,TotFireArea)]
firesDT[,sum(TotFireArea)]
############## All plantations summed ##########
#figuring out area wthin each fire severity category for the whole fire  - might have done this already??
Fire_Sev <- severity_sel %>% filter(FIRE_NUMBE %in% Fire_shortList)
#For every fire, calculate the area (ha) of each severity polygon
Fire_Sev$FireSevArea <- st_area(Fire_Sev)
Fire_SevDT <- as.data.table(Fire_Sev)
FireSevSum <- Fire_SevDT[,round(sum(unclass(FireSevArea)/10000),0),by=c("FIRE_NUMBE","BURN_SEVER")]
setnames(FireSevSum,"V1","FireSev")
#Openings and fire severity
PlantAgeLong <- melt(PlantationsAge[,.(FireID,OPENING_ID,High,Medium,Low,Unburned,TotOpenArea)],
                     id.vars = c("FireID","OPENING_ID","TotOpenArea"))
PlantAgeLong[,PropArea:= value*TotOpenArea]
PlantSevSum <- PlantAgeLong[,round(sum(na.omit(PropArea)),0),by=c("FireID","variable")]
setnames(PlantSevSum,"V1","PlantSev")
PlantFire_Sev <- merge(FireSevSum,PlantSevSum, by.x=c("FIRE_NUMBE","BURN_SEVER"),by.y=c("FireID","variable"))
PlantFire_Sev_Tot <- merge(PlantFire_Sev,firesDTmerge, by="FIRE_NUMBE")
PlantFire_Sev_Tot[,sum(PlantSev),by="FIRE_NUMBE"]
PlantFire_Sev_Tot[,sum(PlantSev)]
PlantFire_Sev_Tot[,sum(PlantSev)]



############### Unioned together ##############
#versus when unioned together (accounting for spatial overlap)
Plant_union <- fread("./Outputs/IndividualFires/PropBurnDat.csv")
Plant_union <- Plant_union %>% filter(FIRE_NUMBE %in% Fire_shortList)
Plant_union[,sum(PlantAreaSev),by="FIRE_NUMBE"]
Plant_union[,sum(na.omit(PlantAreaSev))] #unioned ==22713 - can't be right


PlantFire_Sev_PlUnion <-merge(PlantFire_Sev,Plant_union, by=c("FIRE_NUMBE","BURN_SEVER"))

```



need to reduce the openings to only those that count as plantations. Sheena and Ingrid populated all of them.
```{r}
#
SitePrep[SP_Method=="DISC"|SP_Method=="SHARK"|SP_Method=="",SP_group := "SoilPrep"]


SitePrep[,.N,by=SP_type_meth]
SitePrep[,.N,by=SP_Method]

SitePrep[SP_Method=="PILE"]

ggplot(SitePrep)+
  geom_histogram(aes(x=SP_Date, fill=SP_Method), position="stack")

###### Treatment grouping #####
#treatments that mix the mineral soil:
SoilMix <- c("DISC","MOUND","CHAIN","SHARK","LRIP")
FlattenVeg <- c("","HAND","KNOCK")
TreePiling <- c("RPILE")
BurnTr <- c("SPOT","BROAD","LAND")

SitePrep[,.()]



ResultsSP <- merge(FireDetRes[,.(OPENING_ID, FireID, Num_SitePr=SITE_PRE_5)],SitePrep,by="OPENING_ID")
ResultsSP_PA <- merge(PlantationsAge[FireID %in% Fire_shortList[c(8,9,12)],.(FireID,OPENING_ID,OPENING_ST,
                                                                           SLOPE, ASPECT,High,Low,
                                                                           Medium, Unburned, TotArea,
                                                                           FIRE_YEAR, PropPfire,PlantAge,
                                                                           Spaced,Brushed,SitePrepped,
                                                                           Fertil,Prune)],
                      ResultsSP, by=c("OPENING_ID","FireID"), all=TRUE)




```



