---
title: "PlantationAgeGAM"
author: "Ingrid Farnell"
date: "12/01/2023"
output: html_document
---

This script runs a GAM on variable plantation age and response (dNBR) to further investigate the apparent strong relationship the random forest shows.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#------------------------------ Load libraries---------------------------------#
ls <- c("tidyverse", "data.table") # Data Management and Manipulation
ls <- append(ls, c("vegan", "mgcv")) # analysis

# Install if needed -- then load. 
new.packages <- ls[!(ls %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(ls, library, character.only = TRUE)  # load the required packages
rm(ls, new.packages)


#--------------------------- Load and prep data--------------------------------#
ctg_variables <- c("BEC", "BroadBurn", "Brushed", "DebrisMade", "DebrisPiled", "Fertil", "MechUnk", 
                   "OPENING_ID", "PileBurn", "Prune", "Soil", "Spaced", 
                   "SpotBurn", "WBurn","dNBRCAT")
datPath <- "C:/Users/farne/Documents/Borealis_Ecological_Services/BVRCfire/"     # "./Inputs/"


Chutanli <- fread(paste0(datPath,"G41607dat270.csv"))
Chutanli <- Chutanli %>%
  mutate_at((colnames(Chutanli)[colnames(Chutanli) %in% ctg_variables]), factor) %>%
  dplyr::select(-c("dNBRReSamp")) 
Chutanli[,dNBR := dNBR*1000]
Chutanli[HistoricFires==0 ,HistoricFires:=100]
Chutanli_dist <- dist(Chutanli[,.(x,y)], method = "euclidean")
Chutanli_pcnm <- pcnm(Chutanli_dist)
Chutanli[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Chutanli_pcnm$vectors[,"PCNM1"],
                                            Chutanli_pcnm$vectors[,"PCNM2"],
                                            Chutanli_pcnm$vectors[,"PCNM3"],
                                            Chutanli_pcnm$vectors[,"PCNM4"],
                                            Chutanli_pcnm$vectors[,"PCNM5"],
                                            Chutanli_pcnm$vectors[,"PCNM6"])]

Tezzeron <- fread(paste0(datPath,"G51632dat270.csv"))
Tezzeron <- Tezzeron %>%
  mutate_at((colnames(Tezzeron)[colnames(Tezzeron) %in% ctg_variables]), factor)%>%
  dplyr::select(-c("dNBRReSamp"))
Tezzeron[,dNBR := dNBR*1000]
#Tezzeron[HistoricFires==0 ,Tezzeron:=100] # no historic fires
Tezzeron_dist <- dist(Tezzeron[,.(x,y)], method = "euclidean")
Tezzeron_pcnm <- pcnm(Tezzeron_dist)
Tezzeron[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Tezzeron_pcnm$vectors[,"PCNM1"],
                                            Tezzeron_pcnm$vectors[,"PCNM2"],
                                            Tezzeron_pcnm$vectors[,"PCNM3"],
                                            Tezzeron_pcnm$vectors[,"PCNM4"],
                                            Tezzeron_pcnm$vectors[,"PCNM5"],
                                            Tezzeron_pcnm$vectors[,"PCNM6"])]

Shovel <- fread(paste0(datPath,"R11498dat270.csv"))
Shovel <- Shovel %>%
  mutate_at((colnames(Shovel)[colnames(Shovel) %in% ctg_variables]), factor)%>%
  dplyr::select(-c("dNBRReSamp"))
Shovel[,dNBR := dNBR*1000]
Shovel[HistoricFires==0 ,HistoricFires:=100]
Shovel_dist <- dist(Shovel[,.(x,y)], method = "euclidean")
Shovel_pcnm <- pcnm(Shovel_dist)
Shovel[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Shovel_pcnm$vectors[,"PCNM1"],
                                            Shovel_pcnm$vectors[,"PCNM2"],
                                            Shovel_pcnm$vectors[,"PCNM3"],
                                            Shovel_pcnm$vectors[,"PCNM4"],
                                            Shovel_pcnm$vectors[,"PCNM5"],
                                            Shovel_pcnm$vectors[,"PCNM6"])]

Verdun <- fread(paste0(datPath,"R11796dat270.csv"))
Verdun <- Verdun %>%
  mutate_at((colnames(Verdun)[colnames(Verdun) %in% ctg_variables]), factor)%>%
  dplyr::select(-c("dNBRReSamp"))
Verdun[,dNBR := dNBR*1000]
Verdun[HistoricFires==0 ,HistoricFires:=100]
Verdun_dist <- dist(Verdun[,.(x,y)], method = "euclidean")
Verdun_pcnm <- pcnm(Verdun_dist)
Verdun[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Verdun_pcnm$vectors[,"PCNM1"],
                                            Verdun_pcnm$vectors[,"PCNM2"],
                                            Verdun_pcnm$vectors[,"PCNM3"],
                                            Verdun_pcnm$vectors[,"PCNM4"],
                                            Verdun_pcnm$vectors[,"PCNM5"],
                                            Verdun_pcnm$vectors[,"PCNM6"])]

Island <- fread(paste0(datPath,"R11921dat270.csv"))
Island <- Island %>%
  mutate_at((colnames(Island)[colnames(Island) %in% ctg_variables]), factor)%>%
  dplyr::select(-c("dNBRReSamp"))
Island[,dNBR := dNBR*1000]
Island[HistoricFires==0 ,HistoricFires:=100]
Island_dist <- dist(Island[,.(x,y)], method = "euclidean")
Island_pcnm <- pcnm(Island_dist)
Island[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Island_pcnm$vectors[,"PCNM1"],
                                            Island_pcnm$vectors[,"PCNM2"],
                                            Island_pcnm$vectors[,"PCNM3"],
                                            Island_pcnm$vectors[,"PCNM4"],
                                            Island_pcnm$vectors[,"PCNM5"],
                                            Island_pcnm$vectors[,"PCNM6"])]

Nadina <- fread(paste0(datPath,"R21721dat270.csv"))
Nadina <- Nadina %>%
  mutate_at((colnames(Nadina)[colnames(Nadina) %in% ctg_variables]), factor)%>%
  dplyr::select(-c("dNBRReSamp"))
Nadina[,dNBR := dNBR*1000]
Nadina[HistoricFires==0 ,HistoricFires:=100]
Nadina_dist <- dist(Nadina[,.(x,y)], method = "euclidean")
Nadina_pcnm <- pcnm(Nadina_dist)
Nadina[, c("PCNM1","PCNM2", "PCNM3",
             "PCNM4", "PCNM5","PCNM6") := .(Nadina_pcnm$vectors[,"PCNM1"],
                                            Nadina_pcnm$vectors[,"PCNM2"],
                                            Nadina_pcnm$vectors[,"PCNM3"],
                                            Nadina_pcnm$vectors[,"PCNM4"],
                                            Nadina_pcnm$vectors[,"PCNM5"],
                                            Nadina_pcnm$vectors[,"PCNM6"])]
```

GAM 
- the degree of smoothness is controlled using a penalized regression determined automatically in 'mgcv' using a generalized cross validation routine.
- EDF effective degrees of freedom are related to the smoothing parameter, such that the greater the penalty, the smaller the EDF. Higher EDF imply more complex, wiggly spines (more non-linear).
- gam() sets a default k depending on the number of variables on which the smooth is built
- if the EDF is close to k then the wiggliness of the model is being overly constrained by the default k, and could fit the data better with greater wiggliness (refit the model with a larger k)

- 's()' the smooth function is a 1-dimensional smooth for modeling interactions among variables measured using the same unit and the same scale. There are other smooth functions: te() and ti(), which can both be used to model 2- or n-dimensional interaction surfaces of variables. The function te() is useful when variables are not on the same scale, and when interactions include main effects. The function ti() is best for modelling interaction surfaces that do not include the main effects

The smooth functions have several parameters that can be set to change their behaviour. The most common parameters are:

k: basis dimension - Determines the maximum number of base functions used to build the curve. - Sets the wiggliness of a smooth, in a trade-off with the smoothing parameter. - The k should always be less than the number of unique data points. - The complexity (i.e. non-linearity) of a smooth function in a fitted model is reflected by its effective degrees of freedom (EDF).

bs specifies the type of basis functions. - The default for s() is tp (thin plate regression spline). - The default for te() and ti() is cr (cubic regression spline).
bs = "cr", denotes knot based cubic regression splines

When using te() and ti() basis function, we also need to set the parameter d, which specifies that predictors in the interaction are on the same scale or dimension. - For example, in te(Time, width, height, d=c(1,2)), indicates that width and height are one the same scale, but not Time.

gaussian - normal distribution
link function: identity = model doesn't transform the predictions

```{r echo=FALSE}
#-- Chutanli
gamMod_Chut <- gam(dNBR~s(PlantAge), data = Chutanli, method = "REML") # thin plate regression spline
gamMod_Chutcr <- gam(dNBR~s(PlantAge, bs = "cr", k = 20), data = Chutanli, method = "REML") # cubic regression spline

summary(gamMod_Chut)
gam.check(gamMod_Chut)
gam.check(gamMod_Chutcr)
plot(gamMod_Chut, residuals = TRUE)
plot(gamMod_Chutcr)
plot(gamMod_Chut)

plot <- ggplot(Chutanli, aes(x = PlantAge, y = dNBR)) +
  geom_smooth(method = "gam", 
              formula = y ~ s(x))
plot

AIC(gamMod_Chut, gamMod_Chutcr)
#-- Tezzeron
gamMod_Tez <- gam(dNBR~s(PlantAge), data = Tezzeron, method = "REML")

summary(gamMod_Tez)
gam.check(gamMod_Tez)
plot(gamMod_Tez)


#-- Shovel
gamMod_Sho <- gam(dNBR~s(PlantAge), data = Shovel, method = "REML")

summary(gamMod_Sho)
gam.check(gamMod_Sho)
plot(gamMod_Sho)


#-- Verdun
gamMod_Ver <- gam(dNBR~s(PlantAge), data = Verdun, method = "REML")

summary(gamMod_Ver)
gam.check(gamMod_Ver)
plot(gamMod_Ver)


#-- Island
gamMod_Isl <- gam(dNBR~s(PlantAge), data = Island, method = "REML")

summary(gamMod_Isl)
gam.check(gamMod_Isl)
plot(gamMod_Isl)


#-- Nadina
gamMod_Nad <- gam(dNBR~s(PlantAge), data = Nadina, method = "REML")

summary(gamMod_Nad)
gam.check(gamMod_Nad)
plot(gamMod_Nad)
```

