---
title: "FourFires-Analysis_mlr"
author: "Ingrid Farnell & Alana Clason"
date: "25/02/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script runs ranger using mlr3 package
More info: https://mlr3book.mlr-org.com/introduction.html

Overall data flow:
Step 1: Import data (csvs)
Step 2: Define spatial task
Step 3: Define learner (ranger classification)
Step 4: Separate task into learning (80%) and test (20%) datasets
Step 5: Optimize learner variable importance using spatial cross fold validation
Step 6: Train learner on training data
Step 7: Predict test data using trained learner
Step 8: Evaluate model on test data
Step 9: Visualize 

```{r setup, include=FALSE}
#------------------------------ Load libraries---------------------------------#
ls <- c("tidyverse", "data.table", "magrittr") # Data Management and Manipulation
ls <- append(ls, c("raster","sf")) # geo comp.
ls <- append(ls,c("pdp","mlr3","mlr3spatiotempcv","mlr3verse","ranger"))
ls <- append(ls,c("iml","patchwork","DALEX","DALEXtra"))

# Install if needed -- then load. 
new.packages <- ls[!(ls %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(ls, library, character.only = TRUE)  # load the required packages
rm(ls, new.packages)

```

# STEP 1
Load data as sf objects into a list

```{r}
dat_names <- list.files(path="./Inputs/Shapefiles/", pattern = "cat.gpkg", full.names = TRUE)
dat_list <- purrr::map(dat_names, read_sf)


```

# RF analysis
STEPS 1 - 8

multicorrelated variables are dropped, model is tuned for top pred variables
```{r}
task_names <- c("Chutanli","Island","Nadina","Shovel","Tezzeron","Verdun")
#CV_Conf <- list()
Test_Conf <- list()
TrainImp_dat <- list()
RF_scores <- list()
pd_scores <- list()
effects <- list()
NumCores <- 30

for(i in 1:length(task_names)){
  # Remove correlated covariates or those that are not included in the analysis
  dat <- dat_list[[i]]
  dat <- dat %>% 
    dplyr::select(., -dob, -bui, -dmc, -fwi, -DEMslope, -DEMaspect, 
                     -DFirCov, -FirCov, -PineCov, -SpruceCov)%>% #keep dc, isi for fire behaviour/spread
          mutate_if(sapply(., is.character), as.factor)
  
  task <- as_task_classif_st(dat, target = "dNBRCAT", 
                             id = task_names[i])
  
   #--- Step 1: Define a learner
  # default ranger parameters: https://mlr3learners.mlr-org.com/reference/mlr_learners_regr.ranger.html
  learner <- lrn("classif.ranger",
                importance = "permutation",
                respect.unordered.factors = "order",
                predict_type = "response",
                num.trees = 2000,
                num.threads=NumCores)
  
  #--- Step 2: Split stratified task into test/train data
  set.seed(456)
  split <- partition(task, ratio=0.8, stratify=TRUE)
  task_train <- task$clone()$filter(rows=split$train)
  task_test <- task$clone()$filter(rows=split$test)
  
  #--- Step 3: Feature selection (using nested resampling)
  measure <- msr("classif.ce") #classification error
  fselector <- fs("random_search", batch_size = 30)
  terminator <- trm("evals", n_evals=100)
  inner_resampling <- rsmp("spcv_coords", folds=5)
   
  optFtlrn = AutoFSelector$new(
    learner = learner,
    resampling = inner_resampling,
    measure = measure,
    terminator = terminator,
    fselector = fselector
  )
  
  #--- Step 4: Train final model (applies tuning)
  set.seed(456)
  optFtlrn$train(task_train) 
  
  #--- Step 5: Predict on test data
  optFtlrnPred <- optFtlrn$predict(task_test) 
  optFtlrnPred$score(measure)
  
  #trained model predictions
  #optFtlrnPred$prob
  Test_Conf[[i]] <- optFtlrnPred$confusion
  RF_scores[[i]] <- optFtlrnPred$score()
  
  #--- Step 6: Model visualization
  # Feature importance
  TrainImp_dat[[i]] <- as.data.table(optFtlrn$learner$importance(), keep.rownames = TRUE)
  
  # Feature Imp
  #  Level of importance of the features
  dat_train <- task_train$data()
  x_train <- dat_train[,-c("dNBRCAT")]
  
  model <- Predictor$new(optFtlrn, data = x_train, y = dat_train$dNBRCAT)
  effects[[i]] <- FeatureEffects$new(model)
  model_exp = explain_mlr3(optFtlrn,
                         data = x_train, # provide data without y 
                         y = dat_train$dNBRCAT,
                         colorize = FALSE,
                         type="classification")

  pd_scores[[i]] <- model_profile(model_exp)$agr_profiles
  
} 

#results
RF_scores #trained model predictions on held out data
RF_dat <- data.table(do.call(rbind, Map(cbind, unname(RF_scores), id = task_names)))
setnames(RF_dat, "V1", "CE")
fwrite(RF_dat, "./Outputs/Tables/RF_scores.csv")

names(Test_Conf) <- task_names

TrainImp_dat 

pd_scores 
effects 



```

# Results

```{r}

#my attempt to make that multi-fire graph:
for(i in 1:6){
  TrainImp_dat[[i]][,FireID := task_names[i]]
}
TrainImp_dt <- rbindlist(TrainImp_dat)
setnames(TrainImp_dt, c("Variable","Score","FireID"))
TrainImp_dt[,VarCat:=ifelse(Variable =="PlantAge","Stand structure",
                      ifelse(Variable =="BASAL_AREA","Stand structure",
                       ifelse(Variable =="CROWN_CLOS","Stand structure",
                        ifelse(Variable =="HistoricFires","Previous wildfire",
                         #ifelse(Variable =="DFirCov","Stand structure",
                          #ifelse(Variable =="SpruceCov","Stand structure",
                           #ifelse(Variable =="FirCov","Stand structure",
                            #ifelse(Variable =="decidCov","Stand structure",
                             #ifelse(Variable =="PineCov","Stand structure",
                              ifelse(Variable == "conifCov","Stand structure",
                               ifelse(Variable =="PCNM1","Spatial autocorrelation",
                                ifelse(Variable =="PCNM2","Spatial autocorrelation",
                                 ifelse(Variable =="PCNM3","Spatial autocorrelation",
                                  ifelse(Variable =="PCNM4","Spatial autocorrelation",
                                   ifelse(Variable =="PCNM5","Spatial autocorrelation",
                                    ifelse(Variable =="PCNM6","Spatial autocorrelation",
                                     ifelse(Variable =="DEMhli","Topography",
                                      ifelse(Variable =="DEMtpi","Topography",
                                       ifelse(Variable =="BEC","Climate",
                                       ifelse(Variable =="dc","Fuel moisture",
                                        ifelse(Variable =="isi","Fire behaviour",
                                         ifelse(Variable =="maxT","Fire weather",
                                          ifelse(Variable =="maxRH","Fire weather",
                                           ifelse(Variable =="maxW","Fire weather",
                                            ifelse(Variable =="FireRun","Fire spread",
                                              "Forest Management"))))))))))))))))))))]
TrainImp_dt[,VarNam:=ifelse(Variable =="PlantAge","Plantation Age",
                      ifelse(Variable =="BASAL_AREA","Basal area (m3/ha)",
                       ifelse(Variable =="CROWN_CLOS","Crown closure (%)",
                        ifelse(Variable =="HistoricFires","Time since wildfire",
                         ifelse(Variable =="Soil","Soil disturbance",
                          ifelse(Variable =="BroadBurn","Broadcast burn",
                          ifelse(Variable =="PileBurn","Pile burn",
                            ifelse(Variable =="SpotBurn","Spot burn",
                              ifelse(Variable =="DebrisMade","Debris made",
                                     ifelse(Variable =="DebrisPiled","Debris piled",
                            ifelse(Variable =="decidCov","Deciduous cover (%)",
                             ifelse(Variable =="Fertil","Fertilized",
                              ifelse(Variable == "conifCov","Conifer cover(%)",
                                     ifelse(Variable =="DEMhli","HLI",
                                      ifelse(Variable =="DEMtpi","TPI",
                                       ifelse(Variable =="BEC","BEC",
                                       ifelse(Variable =="dc","drought code (dc)",
                                        ifelse(Variable =="isi","initial spread index (isi)",
                                         ifelse(Variable =="maxT","max temperature",
                                          ifelse(Variable =="minRH","min relative humidity",
                                           ifelse(Variable =="maxW","max wind",
                                            ifelse(Variable =="FireRun","Area burned (%/day)",
                                              Variable))))))))))))))))))))))]
colVar <- unique(TrainImp_dt[,.(Variable,VarCat)])


TrainImp_dt[]
TrainImp_dt[order(FireID, -Score)]


#doesn't work because of ordering
colVar[,VarCatCol:=ifelse(VarCat=="Forest structure", 
                               wes_palette(5,name="Darjeeling1",type="continuous")[][1],
                         ifelse(VarCat=="Spatial", 
                                wes_palette(5,name="Darjeeling1",type="continuous")[][2],
                          ifelse(VarCat=="Climate-Topography",
                                 wes_palette(5,name="Darjeeling1",type="continuous")[][3],
                            ifelse(VarCat=="Fire weather",
                                 wes_palette(5,name="Darjeeling1",type="continuous")[][4],
                                 wes_palette(5,name="Darjeeling1",type="continuous")[][5]))))]
colVar


ggplot(TrainImp_dt, aes(x = Score, y = reorder(VarNam,Score))) +  
  geom_point(aes(colour=FireID), size=3) +
  scale_color_viridis_d(name="Fire",labels=c("Chutanli","Island","Nadina","Shovel","Tezzeron","Verdun"))+
  xlab("Variable importance")+
  ylab("Variables")+
  theme_minimal()+
  theme(strip.text.x = element_text(face="bold"),text=element_text(size=18))
ggsave(filename = "Fig2.jpg",path = "./Outputs/Figures/", device='jpeg', dpi=500, bg="white")

ggplot(TrainImp_dt, aes(x = Score, y = reorder(VarCat,Score))) +  
  geom_point(aes(colour=FireID), size=3) +
  scale_color_viridis_d(name="Fire",labels=c("Chutanli","Island","Nadina","Shovel","Tezzeron","Verdun"))+
  xlab("Variable importance")+
  ylab("Variable groups")+
  theme_minimal()+
  theme(strip.text.x = element_text(face="bold"),text=element_text(size=18))
ggsave(filename = "Fig3.jpg",path = "./Outputs/Figures/", device='jpeg', dpi=500, bg="white")

ggplot(TrainImp_dt, aes(x = Score, y = reorder(VarNam,Score))) +  
  geom_point(aes(colour=FireID), size=3) +
  scale_color_viridis_d(name="Fire",labels=c("Chutanli","Island","Nadina","Shovel","Tezzeron","Verdun"))+
  xlab("Variable importance")+
  ylab("Variable groups")+
  theme_minimal()+
  theme(strip.text.x = element_text(face="bold"),text=element_text(size=12))+
  facet_wrap("FireID")
ggsave(filename = "Fig2_supp.jpg",path = "./Outputs/Figures/", device='jpeg', dpi=300, bg="white")


Silvic_Vars <- c("BroadBurn", "Brushed", "DebrisMade", "DebrisPiled", "Fertil", "PileBurn", 
                  "Prune", "Soil", "Spaced", "SpotBurn")

Silv_Imp_dt <- TrainImp_dt[Variable %in% Silvic_Vars]
ggplot(Silv_Imp_dt, aes(x = Score, y = reorder(Variable,Score))) +  
  geom_point(aes(colour=FireID), size=2) +
  xlab("") +
  xlim(c(-0.01,0.2))+
  ylab("Variable Importance")

num_features <- c("PlantAge")

plot(effects[[1]], features = num_features) 
    #scale_y_continuous("Estimated dNBR") +
    ggtitle("Partial Dependence profiles for selected variables")

plot(effects[[1]]$results$PlantAge$.value ~ effects[[1]]$results$PlantAge$.borders)


PD_plantAge <- list() 
for(i in 1:length(effects)){
  PD_plantAge[[i]] <- data.table(effects[[i]]$results$PlantAge)[,FireID := task_names[i]]
}
PD_plantAge_dt <- rbindlist(PD_plantAge)

ggplot(PD_plantAge_dt)+
  geom_line(aes(x=.borders, y=.value, colour = FireID), size=1.5)+
  facet_wrap(".class")

PD_CrownClos <- list() 
for(i in 1:length(effects)){
  PD_CrownClos[[i]] <- data.table(effects[[i]]$results$CROWN_CLOS)[,FireID := task_names[i]]
}
PD_CrownClos_dt <- rbindlist(PD_CrownClos)

ggplot(PD_CrownClos_dt)+
  geom_line(aes(x=.borders, y=.value, colour = FireID), size=1.5)+
  facet_wrap(".class")

PD_Brushed <- list() 
for(i in 1:length(effects)){
  PD_Brushed[[i]] <- data.table(effects[[i]]$results$Brushed)[,FireID := task_names[i]]
}
PD_Brushed_dt <- rbindlist(PD_Brushed)

ggplot(PD_Brushed_dt)+
  geom_point(aes(x=.borders, y=.value, colour = FireID), size=1.5)+
  facet_wrap(".class")

PD_BroadBurn <- list() 
for(i in 1:length(effects)){
  PD_BroadBurn[[i]] <- data.table(effects[[i]]$results$BroadBurn)[,FireID := task_names[i]]
}
PD_BroadBurn_dt <- rbindlist(PD_BroadBurn)

ggplot(PD_BroadBurn_dt)+
  geom_point(aes(x=.borders, y=.value, colour = FireID), size=1.5)+
  facet_wrap(".class")

PD_Soil <- list() 
for(i in 1:5){
  PD_Soil[[i]] <- data.table(effects[[i]]$results$Soil)[,FireID := task_names[i]]
}
PD_Soil_dt <- rbindlist(PD_Soil)

PD_Brushed <- list() 
for(i in 1:5){
  PD_Brushed[[i]] <- data.table(effects[[i]]$results$Brushed)[,FireID := task_names[i]]
}
PD_Brushed_dt <- rbindlist(PD_Brushed)

PD_DebrisPiled <- list() 
for(i in 1:5){
  PD_DebrisPiled[[i]] <- data.table(effects[[i]]$results$DebrisPiled)[,FireID := task_names[i]]
}
PD_DebrisPiled_dt <- rbindlist(PD_DebrisPiled)

PD_PileBurn <- list() 
for(i in 1:5){
  PD_PileBurn[[i]] <- data.table(effects[[i]]$results$PileBurn)[,FireID := task_names[i]]
}
PD_PileBurn_dt <- rbindlist(PD_PileBurn)


ggplot(PD_Soil_dt)+
  geom_point(aes(x=.borders, y=.value, colour = FireID), size=1.5)+
  facet_wrap(".class")

PD_silv <- rbind(PD_BroadBurn_dt, PD_Soil_dt, PD_Brushed_dt,PD_DebrisPiled_dt,PD_PileBurn_dt)

ggplot(PD_silv)+
  geom_jitter(aes(x=.borders, y=.value, colour=FireID), size=3)+
  #facet_wrap(".class")
  facet_grid(c(".feature",".class"))
  ylim(c(-0.02,0.02))


#probability of unburned:
PD_plantAge_dt
PD_CrownClos_dt


```


